### 2017-01-02 Confirmation/confirmation_outline.Rnw

\documentclass{article}

\begin{document}

\tableofcontents

\section{Bayesian Inference}
\subsection{Bayes Rule}
-General Intro to Bayesian Methodology
-Denominator as an integral
-Often no analytical solutions
\subsection{Exact Bayesian Computation}
-Numerical Integration
-Importance Sampling
-MCMC
\subsection{Variational Bayes}
-KL Divergence as a transformation to distribution space
-VB Derivation
\subsection{The Mean Field Assumption}
-Mean Field Derivation
-Gives optimal family of approximating distributions
-Lots of analytical work
-Very Fast
\subsection{Stochastic Gradient Ascent}
-SGA Derivation
-Step Size
-Flexibility 
-Little analytical work
-Needs a pre-defined approximating distribution family
\subsection{Variance Reduction Techniques}
-Rao Blackwellisation
-Control Variates
-Reparameterisation
\subsection{AR2 model example}

\section{Electricity Load Forecasts}
\subsection{Motivation}
-Real data has five minute updates
-Density is of interest due to rare price spikes, so more interested in the chance load is above a certain point
-Slow convergence of MCMC / Bad scaling to large dimensional models and datasets
\subsection{Exponential Smoothing}
-Bayesian Version
-Difficulty with smoothing parameters in large models
-Infer model structure from MCMC draws where possible
\subsection{Variational Bayes Implementation}
-Posterior Distributions should not change much as more data is observed, so we can keep approximating distribution family constant
-But we might want a model that does change to make updates worthwhile - Markov Switching mechanism to be added?
\subsection{Forecasting}
-Compare to MCMC when possible

\end{document}