### 2017-02-21 Confirmation/confirmation_V2.tex
Revised draft

\documentclass[12pt,a4paper]{article}%
\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{url}
\usepackage{wasysym}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\setcounter{MaxMatrixCols}{30}
\providecommand{\U}[1]{\protect\rule{.1in}{.1in}}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\numberwithin{equation}{section}
\title{Real Time Variational Density Forecasts}
\author{Nathaniel Tomasetti, Catherine Forbes and Anastasios Panagiotelis}
\begin{document}

\maketitle
\tableofcontents
\section{Introduction} \label{sec:Intro}

Electricity prices have been subject to an extensive forecasting literature and are known to be extremely volatile. For the 2011-2015 period in Victoria the mean price per megawatt hour (mWh) was \$42.06, with 99\% of observations less than \$100 per mWh, however on several occasions the price spiked to greater than $10,000\%$ of the mean value. For this reason market participants are increasingly interested in a predictive density forecast, as the probability that the price will rise above a particular threshold in the next time period can be more important than the mean value. To calculate a suitably accurate predictive density the Bayesian methodology is to be adopted in this thesis, as it accounts for both parameter and model uncertainty in a relatively straightforward manner, which can be difficult in frequentist methods. See \citet{Geweke2006} for a review on Bayesian forecasting and \citet{Gneiting2014} for density forecasting in general.

Bayesian methods are computationally complex as they require integration over the set of unknowns, whether parameters or latent stochastic variables. This set is quite large for models used to model short term electricity load and price as there are strong seasonal patterns in the data \citep{Taylor2003}. While analytical integration is typically not feasible to obtain the predictive distribution for such models, however there are a range of numerical and simulation-based techniques available for this purpose, notably Markov Chain Monte Carlo (MCMC) methods. While it is often possible to demonstrate theoretical convergence of MCMC algorithms, computation may be slow for even moderately complex models. This is problematic for when attempting to forecast electricity load in Victoria, when realizations occur at five minute intervals, yhe predictive density for the next period's electricity load may not be available before it is observed.

Alternatives to MCMC include Variational Bayes (VB) methods, where the aim is to replace the unknown true posterior distribution for a vector of unknowns with a parametric approximation. The Variational Bayes approach will be explored in the thesis, to determine if it is a viable alternative computation strategy to MCMC, in the context of modelling and short term forecasting of Victorian electricity load. In particular, we seek to understand the benefits gained in terms of the computational time required to obtain a VB approximate forecast distribution against the loss in statistical accuracy associated with the use of an approximate posterior forecast distribution.

The discussion of MCMC and VB will be continued in Section \ref{sec:BayesInf}, with the methods contrasted in the context of a simple time series model. Section \ref{sec:ElecLoad} will introduce a planned emperical application of VB to Victorian electricity load using an exponential smoothing model and data from the 2012-2015 period. Section \ref{sec:Discussion} outlines the major areas this thesis will make a contribution to and Section \ref{sec:Timeline} outlines the planned timeline of these contributions.

\section{Bayesian Inference} \label{sec:BayesInf}

To facilitate the discussion, consideration is given to models where the unknowns may be summarized in a (finite) $k-$dimensional vector, denoted by $\boldsymbol{\theta}.$ In this setting, given an observed time series, denoted by $y_{1:T}= \{y_{t}, t = 1, \dots T\}$, the Bayesian forecast distribution associated with time $t=T+1$ is characterized by the conditional density
\begin{equation}
\label{predictive}
p(y_{T+1} | y_{1:T}) =\int p(y_{T+1}|y_{1:T},\boldsymbol{\theta}) p(\boldsymbol{\theta}|y_{1:T}) d\boldsymbol{\theta},
\end{equation}
where $p(\boldsymbol{\theta}|y_{1:T})$ denotes the posterior density for $\boldsymbol{\theta}$, given by
\begin{equation}
\label{posterior}
 p(\boldsymbol{\theta} | y_{1:T}) = \frac{p(y_{1:T}|\boldsymbol{\theta})p(\boldsymbol{\theta}}{\int_{\boldsymbol{\theta}}p(y_{1:T}|\boldsymbol{\theta})p(\boldsymbol{\theta}) d\boldsymbol{\theta}}.
\end{equation}

Generally the analytical solution to (\ref{posterior}), and hence to (\ref{predictive}) will be unavailable due to the complex functional form of the distributions involved. 
In this section two alternative methods for computing the desired posterior distribution in (\ref{posterior}),  MCMC and VB, will be reviewed.
In brief, MCMC is used to create a sample from $p(\boldsymbol{\theta} | y_{1:T})$, with any function of $\boldsymbol{\theta}$ that is desired estimated from that sample. In contrast, VB replaces $p(\boldsymbol{\theta} | y_{1:T})$ with an approximation, denoted by $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$, where $\boldsymbol{\lambda}$ is a vector of auxiliary parameters associated with the approximation and may depend on the observations $y_{1:T}$.
For example, if $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ is Gaussian, then $\boldsymbol{\lambda}$ would denote the mean and variance parameters that characeterise this normal distribution for $\boldsymbol{\theta}$.  Variational Bayes aims to choose the distribution $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ such that the Kullback-Leibler divergence from $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ to $p(\boldsymbol{\theta} | y_{1:T})$ is minimised. To obtain the predictive distribution in (\ref{predictive}), $y_{T+1}$ can be treated as an unknown variable in $\boldsymbol{\theta}$ and hence directly computed by either MCMC and VB. Alternatively, once the posterior distribution in (\ref{posterior}) is available, (\ref{predictive}) can be found by simulation.

\subsection{Markov Chain Monte Carlo} \label{sec:MCMC}

There are many types of MCMC algorithms, with arguably the simplest and most commonly used one being the Gibbs sampler. The Gibbs sampler algorithm iteratively samples the components of the $k-$dimensional parameter vector $\boldsymbol{\theta}$ via each of the so-called full conditional distributions as follows,
\begin{align}
&p(\theta_1 | \theta_2, \dots, \theta_k, y_{1:T}) \nonumber \\
&p(\theta_2 | \theta_1, \theta_3, \dots, \theta_k, y_{1:T}) \nonumber \\
&\vdots \nonumber \\
&p(\theta_k | \theta_1, \dots, \theta_{k-1}, y_{1:T}). \nonumber
\end{align}
Under mild regularity conditions (see, e.g., \citet{Tierney1994}) and with enough iterations of the Markov chain that results from the Gibbs sampler, these samples converge in distribution to $p(\boldsymbol{\theta} | y_{1:T})$ and can be used for forecasting and inference about the unknown parameters. The computation time for each iteration and the overall number of iterations required to accruately summarise the posterior distribution is problem specific and typically increases with the number of parameters in the model. On occasion, the full conditional distribution can not be recognised and a Metropolis-Hastings-within-Gibbs step can be used, see \citet{Gilks1995} for details.

For illustration, consider data $y_t$ for $t = 1, \dots, T$ generated independently from a $\mathcal{N}(\mu, \sigma^2)$ distribution with prior distributions given by $\mu | \sigma^2 \sim \mathcal{N}(\gamma, \sigma^2/\tau)$ and $\sigma^2 \sim Inv.Gamma(\mbox{shape} = \alpha, \mbox{scale =}\beta)$. It can be shown that

\begin{align}
\mu | \sigma^2, y_{1:T} &\sim \mathcal{N} \left(\frac{\tau\gamma + T\bar{y}}{\tau + T}, \frac{\sigma^2}{\tau  + T} \right) \label{posmu} \\
\sigma^2 | y_{1:T} &\sim Inv.Gamma  \left( \mbox{shape} = \alpha + T/2, \mbox{scale} = \beta + \frac{1}{2}\sum_{t=1}^{T}y_t^2 + \tau\gamma^2 - \frac{(\tau\gamma + T\bar{y})^2}{\tau + T} \right) \label{possig2} 
\end{align}
where $\bar{y} = 1/T \sum_{t=1}^{T}y_t$. Posterior sampling can proceed by first simulating a draw of $\sigma^2$ from (\ref{possig2}) and conditioning on this draw to simulate $\mu$ from (\ref{posmu}).

\subsection{Variational Bayes} \label{sec:VB}

As discussed in Section \ref{sec:Intro}, a faster (albeit approximate) alternative to MCMC is VB. This approach introduces an approximating distribution $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ with the choice of the functional form $q$ and subsequent auxiliary parameter vector $\boldsymbol{\lambda}$ determined so that the Kullback-Leibler (KL) divergence \citep{Kullback1951} from$q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ the true posterior $p(\boldsymbol{\theta} | y_{1:T})$ is minimised. The KL divergence is defined by
\begin{equation}
\label{KL-def}
KL[q(\boldsymbol{\theta} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta} | y_{1:T}] = \int q(\boldsymbol{\theta} | \boldsymbol{\lambda}) \ln \left( \frac{q(\boldsymbol{\theta} | \boldsymbol{\lambda})}{p(\boldsymbol{\theta} |y_{1:T})}\right) d\boldsymbol{\theta}.
\end{equation}
The KL divergence is a non-negative, asymetric measure of the discrepancy between $p(\boldsymbol{\theta} | y_{1:T})$ and $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$  that will be equal to zero if and only if $p(\boldsymbol{\theta} | y_{1:T}) = q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ almost everywhere \citep{Bishop2006}. Note that $KL[q(\boldsymbol{\theta} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta} | y_{1:T})]$ can be expressed as
\begin{equation}
\label{KL-ELBO}
KL[q(\boldsymbol{\theta} | \boldsymbol{\lambda})\hspace{.1cm}||\hspace{.1cm}p(\boldsymbol{\theta} | y_{1:T})] = \ln(p(y_{1:T})) - \mathcal{L}(q, y_{1:T})
\end{equation}
where $\mathcal{L}(q, y_{1:T})$ is referred to as the Evidence Lower Bound (ELBO), as it provides a lower bound on the unknown constant $\ln(p(y_{1:T}))$.  $\mathcal{L}(q, y_{1:T})$ is defined by
\begin{equation}
\label{ELBO}
\mathcal{L}(q, y_{1:T}) = \int_{\boldsymbol{\theta}} q(\boldsymbol{\theta} | \boldsymbol{\lambda}) \ln (p(y_{1:T},\boldsymbol{\theta})) d\boldsymbol{\theta} -  \int_{\boldsymbol{\theta}} q(\boldsymbol{\theta} | \boldsymbol{\lambda}) \ln (q(\boldsymbol{\theta} | \boldsymbol{\lambda})) d\boldsymbol{\theta}.
\end{equation}

Since $\ln(p(y_{1:T}))$ is constant with respect to $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$, maximising (\ref{ELBO}) with respect to $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ is equivalent to minimising (\ref{KL-def}). Maximising the ELBO is much more convenient than minimising the KL Divergence, as will be shown in the following sections in the context of the two major implementations of Varitaional Bayes, namely Mean Field Variational Bayes and Stochastic Variational Bayes. Each of these implementations take advantage of the functional form of the ELBO, and offer computationally efficient algorithms to find a $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ that is optimal within a particular class of distributions. 
of distributions.

\subsubsection{Mean Field Variational Bayes} \label{sec:mfvb}

Mean Field Variational Bayes (MFVB) has origins in physics \citep{Chandler1987} and has since developed a large literature in machine learning, notably developed in \citet{Jordan1999} and \citet{Beal2003}. MFVB restricts the class of distributions from which to select $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ to the set of factorisable distributions,
\begin{equation}
\label{mf1}
q(\boldsymbol{\theta} | \boldsymbol{\lambda}) = \prod_{i=1}^k q_i(\theta_i | \boldsymbol{\lambda_i}).
\end{equation}

Each of the $k$ components $\theta_i$ may be a vector, but in most implementations are scalars. Each $\theta_i$ has an associated vector $\boldsymbol{\lambda_i}$ which may be of a different dimension to $\theta_i$. $\boldsymbol{\lambda_i}$ is an auxillary parameter vector for the relevant factor $q_i$, which will be used in this section as shorthand notation for the distribution $q_i(\theta_i |\boldsymbol{\lambda_i})$.

MFVB is widely used as it greatly simplifies maximisation of the ELBO, and has a straight-forward implementation in exponential family models \citep{Wainwright2008}.  Maximising the ELBO with respect to $q_i$ is as analytically involved as deriving the conditional distributions used in Gibbs based MCMC schemes, but the resulting algorithm to maximise the ELBO is much simpler computationally than sampling from a Gibbs MCMC scheme. MFVB factorises (\ref{ELBO}) into a product of $k$ seperate components that each contain only one $q_i$, and then applies an algorithm to maximise each of these factors with respect the associated $q_i$. Using the notation $q_{\setminus i} = \prod_{j\neq i}q_j$, \citet{Attias1999} shows that each component is maximised by
\begin{equation}
\label{mf2}
q_i(\theta_i |\boldsymbol{\lambda_i}) \propto\exp( \mathbb{E}_{q_{\setminus i}} [\ln(p(y_{1:T},\boldsymbol{\theta}))]).
\end{equation}

We illustrate the implementation of MFVB with the model considered in Section (\ref{sec:MCMC}), where $y_t$ for $t = 1, \dots, T$ is generated independently from a $\mathcal{N}(\mu, \sigma^2)$ distribution, and prior distributions are given by $\mu \sim \mathcal{N}(\gamma, \sigma^2 / \tau)$ and $\sigma^2 \sim Inv.Gamma(\mbox{shape} = \alpha, \mbox{scale =}\beta)$. It can be shown that
\begin{equation}
\label{mf3}
q(\mu | \boldsymbol{\lambda}_1) \propto \exp \left\{ \frac{-(T\tau + \mathbb{E}_{q(\sigma^2)}(\sigma^{-2}))}{2\tau\mathbb{E}_{q(\sigma^2)}(\sigma^{-2})} \left( \mu - \frac{\mathbb{E}_{q(\sigma^2)}(\sigma^{-2})\gamma + \lambda \sum_{t=1}^{T} y_t}{T \tau + \mathbb{E}_{q(\sigma^2 )}(\sigma^{-2})} \right)^2 \right\}
\end{equation}
and
\begin{equation}
\label{mf4}
q(\sigma^2 | \boldsymbol{\lambda}_2) \propto \sigma^{-2(T/2 + \alpha + 1)} \exp \left\{ \frac{ -1/2(\sum_{t=1}^{T}y_t^2 + n\mathbb{E}_{q(\mu)}(\mu^2) - 2\sum_{t=1}^{T} y_t \mathbb{E}_{q(\mu)}(\mu)) - \beta}{\sigma^2} \right\}.
\end{equation}

As the prior distributions used were conjugate to the likelihood, the Variational Bayes optimal distributions for $q(\mu |  \boldsymbol{\lambda}_1)$ and $q(\sigma^2 | \boldsymbol{\lambda}_2)$ are the same family as the prior distributions, and are denoted by $q(\mu | \boldsymbol{\lambda}_1) \sim \mathcal{N}(\tilde{\gamma}, \tilde{\tau})$ and $q(\sigma^2 | \boldsymbol{\lambda}_2) \sim Inv.Gamma(\tilde{\alpha}, \tilde{\beta})$. The vector $\boldsymbol{\lambda} = (\tilde{\gamma}, \tilde{\tau}, \tilde{\alpha}, \tilde{\beta})'$ is given by
\begin{align}
\tilde{\gamma} &= \frac{\mathbb{E}_{q(\sigma^2)}(\sigma^{-2})\gamma + \lambda \sum_{t=1}^{T} y_t}{T \tau + \mathbb{E}_{q(\sigma^2 )}(\sigma^{-2})}  \label{mf5} \\ 
\tilde{\tau} &= \frac{\tau\mathbb{E}_{q(\sigma^2)}(\sigma^{-2})}{(T\tau + \mathbb{E}_{q(\sigma^2)}(\sigma^{-2}))} \label{mf6} \\
\tilde{\alpha} &= T/2 + \alpha  \label{mf7} \\
\tilde{\beta} &= 1/2\left(\sum_{t=1}^{T}y_t^2 + n\mathbb{E}_{q(\mu)}(\mu^2) - 2\sum_{t=1}^{T} y_t \mathbb{E}_{q(\mu)}(\mu))\right) + \beta. \label{mf8}
\end{align}.

The known values of the expectations in (\ref{mf5}) - (\ref{mf8}) for these distributions can be substituted into these expressions to derive the set of mean field equations:

\begin{align}
\tilde{\gamma} &= \frac{\tilde{\alpha} / \tilde{\beta} \gamma + \tau \sum_{t=1}^{T} y_t} {T \lambda + \tilde{\alpha} / \tilde{\beta}} \label{mf9} \\ 
\tilde{\tau} &= \frac{T \tilde{\alpha} / \tilde{\beta}}{T \tau + \tilde{\alpha} / \tilde{\beta}} \label{mf10} \\
\tilde{\beta} &= \frac{1}{2} \left(\sum_{t=1}^{T} y_t^2 + T(\tilde{\gamma}^2 + \tilde{\tau}) - 2 \sum_{t=1}^{T} y_t \tilde{\gamma} \right) + \beta. \label{mf11}
\end{align}
$\tilde{\alpha}$ is omitted from the mean field equations as its value does not depend on other $\boldsymbol{\lambda}$ parameters. (\ref{mf9}) - (\ref{mf11}) contain a circular dependence in these equations an algorithm that iteratively cycles through these equations until the change in the ELBO is less than some pre-specified constant $\epsilon$. This algorithm is known as a coordiant ascent algorithm and is described by Algorithm (\ref{alg:algorithm1}), at each step it will fix all but one parameter and select the value of the omitted parameter which maximises the ELBO.

The set of mean field equations are only obtainable in the event that a prior that is conjugate to the model likelihood is used, restricting MFVB to exponential family models. In this case the optimal distributional family $q_i$ is the same as the prior, analagous to the relationship between a prior-posterior conjugate pair. In the event that a non-conjugate prior is used, the form of the distribution in (\ref{mf2}) is unrecognisable, a further approximation may be used by substituting in another distribution to replace the problematic $q_i(\theta_i |\boldsymbol{\lambda_i}).$ A common option is the use of a Gaussian distribution through either a Laplace approximation \citep{Friston2006}, a Delta Method inspired transform \citet{Wang2013}, though the only requirement is that the substitute distribution has moments that are tractable functions of its parameters to be used for other parameters in (\ref{mf2}). This secondary level of approximations is analogous to the requirement of a Metropolis-Hastings-within-Gibbs step in MCMC to handle unrecognisable full conditional distributions.

\vspace{5mm}
\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Log Joint Density}
 \KwResult{Mean Field Approximation}
 Use (\ref{mf3}) to match $q(\mu|\boldsymbol{\lambda}_1)$ to a Gaussian distributon and $q(\sigma^2 | \boldsymbol{\lambda}_2)$ to an Inverse Gamma distribution\;
 Anaytically derive the set of mean field equations in \ref{mf9}) - (\ref{mf11})\;
 Initialise $\boldsymbol{\lambda}^{(1)}$ randomly.\;
 Evaluate $\mathcal{L}(q, y_{1:T})^{(1)}$ using $\boldsymbol{\lambda}^{(1)}$.\;
 Set $\tilde{\alpha}$ using \ref{mf7}.\;
 \While{$\mathcal{L}(q, y_{1:T})^{(m)} - \mathcal{L}(q, y_{1:T})^{(m-1)} > \epsilon$}{
  Set $\tilde{\gamma}^{(m)}$ using \ref{mf9}.\;
  Set $\tilde{\tau}^{(m)}$ using \ref{mf10}.\;
  Set $\tilde{\beta}^{(m)}$ using \ref{m11}.\;
  Evaluate $\mathcal{L}(q, y_{1:T})^{(m)}$.\;
  Set $m = m + 1$.\;
 }
 \caption{Coordinate Ascent for MFVB}
  \label{alg:algorithm1}
\end{algorithm}

\subsubsection{Stochastic Variational Bayes}

The requirement for an exponential family model, as well as restricting the approximating distribution to a factorisable family may be unsatisfactory in many applications where the dependence between parameters is important, as in the AR(2) model where Figure \ref{fig:MCMCplot} shows strong dependence between $\phi_1$ and $\phi_2$. Without factorisation the simple method to maximise the ELBO as in MFVB is unavailable. \citet{Paisley2012} and \citet{Ranganath2014} have adapted a gradient ascent algorithm for use in Variational Bayes, resulting in what is known as Stochastic Variational Bayes (SVB).

The algorithm for SVB iteratively takes the  derivative of $\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}), y_{1:T})$ in (\ref{ELBO}) with respect to $\boldsymbol{\lambda}$ and the following updating step is applied until the ELBO converges to a maximum.

\begin{equation}
\label{SGA1}
\boldsymbol{\lambda}^{(m+1)} = \boldsymbol{\lambda}^{(m)} + \rho^{(m)} \nabla_{\boldsymbol{\lambda}} \mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T}),
\end{equation}

where $\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ is the vector of partial derivatives of $\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T})$ with respect to each element of $\boldsymbol{\lambda}$. This update requires some initial values $\boldsymbol{\lambda}^{(0)}$ and a sequence $\rho^{(m)}, m = 1, 2, \dots$ known as the learning rate. If $\rho^{(m)}$ is chosen to satisfy the following conditions the algorithm is guaranteed to converge to a local maximum \citep{Robbins1951}.

\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}

Whilst a global maximum is desired, Variational Bayes theory does not offer any insight to the shape of the ELBO so determining if a local maximum is a global maximum is difficult. One option to alleiviate this problem is to start the algorithm at a range of initial values choose the converged value with the highest maximised ELBO and hence lowest KL divergence to the true posterior.

SVB can find only the optimal values for $\boldsymbol{\lambda}$ for a user specified distributional family $q$, and to run SVB many choices of $q$ may be used and one selected on the basis of the highest ELBO. $q$ is restricted to the family of distributions that satisfy the condition that the order of differentation of the ELBO with respect to $\boldsymbol{\lambda}$ and integration with respect to $\boldsymbol{\theta}$ are interchangable, and SVB can only be applied to models where the log likelihood for $y_{1:T}$ is able to be evaluated. Given these conditions \citet{Ranganath2014} showed that a Monte Carlo estimate of the derivative of the ELBO can be taken by

\begin{equation}
\label{SGA2}
\nabla_{\lambda}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_s | \boldsymbol{\lambda}^{(m)}))] (\ln (p(y_{1:T}, \boldsymbol{\theta}_s)) - \ln(q(\boldsymbol{\theta}_s | \boldsymbol{\lambda}^{(m)})))
\end{equation}

where $s = 1, \dots, S$ indicates simulations from $q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)})$.

\citet{Duchi2011} introduced the AdaGrad algorithm which can be implemented within SVB to control $\rho^{(m)}$. AdaGrad allows each $\lambda_i$ to have an independent $\rho^{(m)}_i$ that is inversely proportional to the gradient, so $\lambda$ takes bigger steps in flat regions and smaller steps in steep regions.

Let

\begin{equation}
\label{SGA3}
G_i^{(m)} = \sum_{j = 1}^{m} \left(\nabla_{\lambda_i}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(j)}), y_{1:T})\right)^2,
\end{equation}

then each component's learning rate is defined as

\begin{equation}
\label{SGA4}
\rho^{(m)}_i = \eta \left(G_i^{(m)}\right)^{-1/2}
\end{equation}

for some tuning parameter $\eta$.

The resulting Stochastic Gradient Ascent algorithm proceeds below with a $p$ dimensional $\boldsymbol{\lambda}$ vector.

\vspace{2mm}

\begin{algorithm}[H]
 \SetKwInOut{Input}{Input}
 \Input{Log Joint Density, Approximation family q}
 \KwResult{Variational Approximation}
 Initialise $\boldsymbol{\lambda}^{(1)}$\;
 Evaluate $\mathcal{L}(q, y_{1:T})^{(1)}$ using $\boldsymbol{\lambda}^{(1)}$.\;
 \While{$\mathcal{L}(q, y_{1:T})^{(m)} - \mathcal{L}(q, y_{1:T})^{(m-1)} > \epsilon$}{
  Simulate $\boldsymbol{\theta}^s$ for $s = 1, \dots S$ from $q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)})$\;
  \For{$i =  1$ \KwTo $p$}{
      Calculate $\nabla_{\lambda_i}$ from (\ref{SGA2})\;
      Update $G_i^{(m)}$ and $\rho^{(m)}_i$ from (\ref{SGA3}) and (\ref{SGA4})\;
      }
  Update $\boldsymbol{\lambda}^{(m+1)}$ from (\ref{SGA1})\;
  Evaluate $\mathcal{L}(q, y_{1:T})^{(m)}$\;
  Set $m = m + 1$\;
 }
 \caption{Stochastic Gradient Ascent for SVB}
  \label{alg:algorithm2}
\end{algorithm}

\subsection{Vine Copulas}

Armed with an algorithm to find the Variational Bayes optimal parameters $\boldsymbol{\lambda}$ for a distribution $q$, there is a requirement to have a distribution $q$ that is itself optimal. Copulas are a flexible tool for constructing these distributions, as they allow the dependence structure between parameters to be fit independently from the marginal distributions. \citet{Sklar1959} proves that any joint probability distribution can be written as the product of the marginals and a copula function,

\begin{equation}
\label{vc1}
p(\theta_1, \dots, \theta_k) = p(\theta_1) \dots p(\theta_k) c(P(\theta_1), \dots, P(\theta_k))
\end{equation}

where $p(\theta)$ is a pdf, $P(\theta)$ is a cdf, and $c(\cdot)$ is a copula.

This idea is extended by the vine copula, a technique to factorise a high-dimensional copula into a set of bivariate copulas in a similar way to the factorisation of a joint distribution into a marginal and conditional distribution, see \citet{Joe2014} and references within. An example of this factorisation is

\begin{equation}
\label{vc2}
c(P(\theta_1), P(\theta_2), P(\theta_3)) = c(P(\theta_1), P(\theta_2)) \cdot c(P(\theta_1), P(\theta_3)) \cdot c(P(\theta_2), P(\theta_3) | P(\theta_1)).
\end{equation}

A copula for a $k$ dimensional $\theta$ can be replaced with $k(k-1)/2$ bivariate copulas arranged in a vine copula, greatly increasing the amount of flexibility available to fit a distribution. However factorising the copula and choosing a family for each bivariate copula takes exponential time in the number of parameters, so without detailed problem specific information the application of a vine copula to Stochastic Variational Bayes is difficult. \citet{Tran2015} implements Variational Bayes with a vine copula in what they call Copula Variational Bayes, but the dimension of $\boldsymbol{\theta}$ was small in that case allowing an exhaustive search for the vine copula that maximised the ELBO.

Given a sample of $\boldsymbol{\theta}$, \citet{Dissmann2013} provide an algorithm to select the best vine factorisation by maximising Kendall's Tau, and the best family for each bivariate copula by minimising an information criterion such as AIC. In the real-time forecasting models used for electricity load demand, MCMC samples of the posterior distribution are available, though often immediately out of date. The approach used in this thesis is to run an MCMC algorithm, thin the resulting draws until they are effectively independent, choose an optimal family for each marginal distribution by AIC, then run (Vine copula algorithm here) to infer an optimal vine copula. This results in an approximating distribution $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ that can then be used in an SVB algorithm to update $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ as more observations become available. MCMC can be ran simultaenously and the resulting sample can be used to check if the distribution $q$ should be changed.

\subsection{Applying Stochastic Variational Bayes}

To illustrate issues involved, consider {\bf the second order} auto-regressive time series {model, denoted by AR(2) and described by}
% no need to say more than this here.
% But actually, I think you should just do the very simple Normal Inverse Gamma model here, since that has the parallel with VB below. Then in your example later on, you can do the full AR(2) example. I think this will help a lot.

\begin{equation}
\label{AR2}
y_t = \rho_1 y_{t-1} + \rho_2 y_{t-2} + \epsilon_t
\end{equation}

where $\epsilon_t \sim \mathcal{N}(0, \sigma^2)$. The likelihood of parameters {\bf $\theta = (\rho_1, \rho_2, \sigma^2)'$, that is, the density of the observations $y_{1:T}$, given $\theta$, viewed as a function of $\theta$,)} is given by

\begin{align}
\label{likelihood}
L(\theta | y_{1:T}) &= p(y_1, y_2 | \theta) \prod_{t=3}^{T}p(y_t | y_{1:t-1}, \theta) \nonumber \\
&= \frac{1}{(2\pi)} |\Sigma|^{-1/2} \sigma^{-(T-2)} \exp \left\{ \frac{-1}{2} \left( \textbf{y}' \Sigma^{-1} \textbf{y} + \frac{1}{\sigma^2} \sum_{t=3}^{T}(y_t - \phi_1 y_{t-1} - \phi_2 y_{t-2})^2 \right) \right\}
\end{align}

where

\begin{equation}
\Sigma = \left[ \begin{array}{cc} \gamma(0) & \gamma(1) \\ \gamma(1) & \gamma(0) \end{array} \right],\
\end{equation}
% Changed the matrices above to align properly
$\gamma(k)$ is the $k-th$ autocorrelation and $\textbf{y} = (y_1, y_2)'$.
% You also need to mention what $p(y_1, y_2 | \theta)$ is, and perhaps something about how $p(y_t | y_{1:t-1}, \theta)$ is available from the assumptions regarding the distribution of $\epsilon_t$.
% Also, I don't like the use of vector notation (for $y=(y_1,y_2)$ )and subscripts for the remaining $y$'s in the likelihood. You can just break the equation over two lines if needed. 
% Finally, perhaps you could say something about where $\gamma_0$ and $\gamma_1$ come from, or at least provide a reference for the reader to go and find out more about this model so you don't have to go into too much detail here.

{\bf We assume that the prior distribution, is given by }
% remember Tas's comment about stating that the joint prior is assumed to be the product of independent marginal priors

\begin{align}
p(\sigma^2) &\propto \sigma^{-2} \nonumber \\
p(\rho_1, \rho_2) &\propto \mathbb{I}(\rho_2 > -1)\mathbb{I}(\rho_2 < 1 + \rho_1) \mathbb{I}(\rho_2 < 1 - \rho_1) \nonumber
\end{align}

are used, where $\mathbb{I}$ is the indicator and $p(\rho_1, \rho_2)$ defines the AR(2) stationary region.

{\bf Upon examination of the form of the product of the likelihood function in (\ref{likelihood}) and the prior density in (),}
	%need to be able to reference the prior too
{\bf it can be seen that the full} conditional distribution $p(\sigma^2 | y, \phi_1, \phi_2)$ is {\bf an} Inverse Gamma {\bf distribution},
% can you say which one? Be sure it is clear how to interpret the parameters of this distribution too.
 while the conditionals for both $\phi$ parameters are intractable and sampled using a Metropolis-Hastings step 
% Since you are actually using an M-H step, you should have told the reader what this is back up when you introduced the Gibbs sampler.
with truncated {\bf bivariate} %?
normal random walk {proposal distribution}. The truncation ensures that draws {\bf of $(\rho_1,\rho_2)$} %just to be clear
are only taken from the AR(2) stationary region.

%Any graph that you put in to the document needs to be clearly explained. Why discard any? What does 'thinned' mean, and why did you do it? etc.... Better to put the details of the MCMC in a paragraph, and then point to the figure, with only the essential details left in the caption. 
% Also, ideally you would be able to put the Greek symbols in the margins of the graphs, rather than 'Sigmasquared', for example...

\begin{figure}[h]
\centering
\includegraphics[scale = 0.6]{MCMC.png}
\caption{MCMC draws for the AR(2) model with $T = 150, \phi_1 = 0.75, \phi_2 = 0.2$ and $\sigma^2 = 1$. Approximately 24\% of both $\phi$ parameters were accepted by the Metropolis Hastings step, then the first 10,000 draws were discarded and the remainder thinned to 1 in 200 to make dependence between draws negligible.}
\label{fig:MCMCplot}
\end{figure}

Using the approach described in section 2.3, an application of VC algorirhtm to the MCMC draws obtained in section 2.1 found the following distribution to be optimal by AIC.

\begin{itemize}
\item $q(\phi_1 | \boldsymbol{\lambda}_1)$ - Gaussian marginal
\item $q(\phi_2| \boldsymbol{\lambda}_2)$ - Gaussian marginal
\item $q(\sigma^2| \boldsymbol{\lambda}_3)$ - Inverse Gamma marginal
\item $c(Q(\phi_1), Q(\phi_2) | \boldsymbol{\lambda}_4)$ - Gaussian copula
\item $c(Q(\phi_1), Q(\sigma^2))$ - Independent
\item $c(Q(\phi_2), Q(\sigma^2) | Q(\phi_1))$ - Independent
\end{itemize}

Figure \ref{fig:VBfit} shows that Stochastic Variational Bayes (blue) fit the MCMC (red) sample very closely.

\begin{figure}[h]
\centering
\includegraphics[width=0.7\linewidth,height=\textheight,keepaspectratio]{VBfit.png}
\caption{The fit of an SVB algorithm (red) compared to MCMC (red) for the AR(2) model described in section 2.1. SVB resulted in an almost identical posterior to MCMC, requiring around one hundredth of the runtime for MCMC.}
\label{fig:VBfit}
\end{figure}

\section{Electricity Load Forecasts} \label{sec:ElecLoad}
\subsection{Motivation}

The literature for electricity forecasting for both load and price is wide, and contains models based on many different fields, such as game theory, time-series modelling and neural networks, see \citet{Weron2014} for a recent review. Density forecasting for short term electricity load has had less attention, with \citet{Fan2012} and \citet{He2016} being notable examples, which used boot strapping and quantile regression respectively to generate forecast densities. Both of these approaches create forecast densities that are conditioned on the model and estimated parameters being correct, while the Bayesian approach used in this thesis can average over parameter and model uncertainty leading to a more accurate forecast density. This thesis also aims to re-estimate the model after each data point is observed to fully take advantage of any extra information on changes in parameters and thus further increase the accuracy of the forecast density.

Half hourly electricity load data for Victoria, Australia has been collected for the 2011-2015 period, and is plotted in Figure \ref{fig:loadplot}. The time series displays strong seasonal characteristics, with yearly, day of week, and time of day patterns. Electricity load is strongly dependent on temperature, with high load experienced on both very cold and very hot days. In summer periods, load is often low but shows extreme volatility, while load slowly rises then falls in winter and displays significantly less volatility.

The electricity market in Victoria operates by market participants sending electricity bid and offers for each five minute period to  AEMO, a centralised market operator. AEMO then determines the dispatch price and load for that five minute period, and determines the final spot price participants pay for each half-hour period by averaging the six five minute dispath prices in that half-hour. Market participants are allowed to revise bids and offers at any point before dispatch. The electricity supply curve is strongly hockey stick shaped, as renewables and coal generators are able to provide a large load cheaply but if demand exceeds this level the price spikes as other generators must be brought online expensively. These characteristics of the electricity market leads to a large amount of volatility in prices demonstrated by Figure \ref{fig:rrpplot}. The seasonal effects of the load data are also present in prices, but these effects are dominated by irregular spikes as high as \$9,974/mWh, compared to a four year mean of \$42/mWh.

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth,height=\textheight,keepaspectratio]{load_timeseries.png}
\caption{Half hourly load data measured in megawatt hours for Victoria from January 1 2011 to December 31 2015. The yearly pattern is characterised by a low median load with regular spikes in summer and a more consistent curve in winter.}
\label{fig:loadplot}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth,height=\textheight,keepaspectratio]{RRP_timeseries.png}
\caption{Half hourly price data for Victoria from January 1 2011 to December 31 2015. While the price has time of week and time of day seasonal effects these features are dominated by irregular spikes.}
\label{fig:rrpplot}
\end{figure}

\subsection{Exponential Smoothing}

\citet{Taylor2003} explores the seasonal properties of minute-by-minute electricity demand in the United Kingdom and introduces a double seasonal Holt-Winters exponential smoothing model with daily and weekly effects described by (\ref{ds-hw1})-(\ref{ds-hw4}). Taylor omits yearly effects as their inclusion requires a large number of latent states to be estimated. This model is later verified in \citet{Taylor2008} as superior to other common time-series models for very short-term load forecasting.

\begin{align}
y_t &= l_{t-1} + d_{t-m_1} + w_{t-m_2} + e_t \label{ds-hw1} \\
l_t &= \alpha (y_t - d_{t-m_1} - w_{t-m_2}) + (1 - \alpha)l_{t-1} \label{ds-hw2}\\
d_t &= \delta (y_t - l_{t-1} - w_{t-m_2}) + (1 - \delta)d_{t-m_1} \label{ds-hw3} \\
w_t &= \omega (y_t - l_{t-1} - d_{t-m_1}) + (1 - \omega)w_{t-m_2} \label{ds-hw4}
\end{align}

where $m_1$ and $m_2$ are the lengths of the daily and weekly cycle, and the smoothing parameters $\alpha, \delta, \omega$ are also restricted to to lie in $(0, 1)^3$. This can be rewritten as a single source of error state-space model \citep{Snyder1985},

\begin{align}
y_t &= l_{t-1} + d_{t-m_1} + w_{t-m_2} + e_t \label{ds-hw-ssoe1} \\
l_t &= l_{t-1} + \alpha e_t \label{ds-hw-ssoe2} \\
d_t &= d_{t-m_1} + \delta e_t \label{ds-hw-ssoe3} \\
w_t &= w_{t-m_2} + \omega e_t \label{ds-hw-ssoe4}. 
\end{align}

The unknown parameters are $\boldsymbol{\theta} = (\alpha, \delta, \omega)', \sigma^2$ and $\boldsymbol{b}_0 = (l_0, d_0, \dots, d_{-(m_1 - 1)}, w_0, \dots, w_{-(m_2 - 1)})'$ a $k = m_1 + m_2 + 1$ length vector of the initial states of the latent variables.

\citet{Forbes2000} provides a transformation to the model that allows a straight-forward Bayesian analysis in small samples, noting that the state space model can be expressed as

\begin{align}
y_t &= \textbf{x}' \boldsymbol{b}_t + e_t \label{ss_me} \\
\textbf{b}_t &= \textbf{T} \textbf{b}_{t-1} + \tilde{\boldsymbol{\theta}} e_t \label{ss_te}.
\end{align}

Using $D = T - \tilde{\boldsymbol{\theta}} \textbf{x}'$, the transition equation in (\ref{ss_te}) can be written as

\begin{equation}
b_t = D b_{t-1} + \tilde{\boldsymbol{\theta}} y_t \label{ss_te_2}.
\end{equation}

Setting $\bar{b}_0$ = 0, the recursion in (\ref{ss_te_2}) allows the transformed observations $\tilde{y}_t$ to be obtained via

\begin{equation}
\tilde{y}_t = y_t - x'\bar{b}_{t-1} \label{ss_rp_y}.
\end{equation}

Furthermore, transformed regression variables $\tilde{\textbf{x}}_t$ can be obtained from $\tilde{\textbf{x}}_t = D\tilde{\textbf{x}}_{t-1}$ with $\tilde{\textbf{x}}_0 = \textbf{x}$. Collecting these as $\widetilde{X}' = (\tilde{\textbf{x}}_1, \tilde{\textbf{x}}_2, \dots, \tilde{\textbf{x}}_T)$ and $\widetilde{Y}' = (\tilde{y}_1, \tilde{y}_2, \dots, \tilde{y}_T)$, and assuming the errors $e_t$ are Gaussian, the model likelihood is

\begin{equation}
L(\boldsymbol{\theta}, \sigma^2, \boldsymbol{b}_0 | y_{1:T}) \propto \sigma^{-T} \exp \left\{ \frac{-1}{2\sigma^2}(\widetilde{Y} - \widetilde{X}\boldsymbol{b}_0)'(\widetilde{Y} - \widetilde{X}\boldsymbol{b}_0) \right\}.
\end{equation}

Hence once conditioned on a draw of $\boldsymbol{\theta}$, the remaining unknown parameters can be sampled as a Bayesian linear regression problem.

\citet{Forbes2000} provides the marginal distribution of $\theta$,

\begin{equation}
\label{exp-sm-marginal}
p(\boldsymbol{\theta} | y_{1:T}) \propto \left| \widetilde{X}' \widetilde{X} \right|^{-1/2} \tilde{s}^{-(T-k)} p(\theta),
\end{equation}

where $\tilde{s}^2 = (\widetilde{Y} - \widetilde{X}\hat{\boldsymbol{b}}_0)'(\widetilde{Y} - \widetilde{X}\hat{\boldsymbol{b}}_0) / (T - (m_1 + m_2 + 1))$ is the sum of squared errors for $\hat{\boldsymbol{b}}_0 = (\widetilde{X}'\widetilde{X})^{-1}\widetilde{X}'\widetilde{Y}$.
With the seasonality parameterisation in (\ref{ds-hw-ssoe1}) - (\ref{ds-hw-ssoe4}), the matrix $\widetilde{X}' \widetilde{X}$ is singular, but restricting the latent variables via

\begin{align}
d_t &= - \sum_{i=1}^{m_1-1} d_{t-i} \\
w_t &= - \sum_{i=1}^{m_2-1} w_{t-i}
\end{align}

avoids this problem. The resulting model is described by

\begin{align}
y_t &= l_{t-1} - \sum_{i = 1}^{m_1 - 1}d_{t-i} - \sum_{i = 1}^{m_2 - 1}w_{t-i} + e_t \label{ds-hw-rp1} \\
l_t &= l_{t-1} + \alpha e_t \label{ds-hw-rp2} \\
d_t &= - \sum_{i = 1}^{m_1 - 1}d_{t-i} + \delta e_t \label{ds-hw-rp3} \\
w_t &= - \sum_{i = 1}^{m_2 - 1}w_{t-i} + \omega e_t \label{ds-hw-rp4}.
\end{align}

and $\boldsymbol{b}_0 = (l_0, d_0, \dots, d_{-(m_1 - 2)}, w_, \dots, w_{-(m_2 - 2)})'$ is a length $k = m_1 + m_2 - 1$ vector.

\citet{Forbes2000} recommends numerical integration of the marginal posterior density of $\theta$ in (\ref{exp-sm-marginal}), however the problems with using this method in our model are two-fold: repeating a three dimensional numerical integrion each time a new data point is observed within the five minute window is not feasible, and the $T$ in the exponent makes most evalutions of $p(\theta | y_{1:T})$ computationally zero when $T$ is large, as is the case with high frequency electricity load data. An efficient MCMC algorithm using the conditional distribution of $p(\theta | y_{1:T}, \sigma^2, b_0)$ may be required to sample the full posterior distribution $p(\theta, \sigma^2, b_0 | y_{1:T}$ and obtain an approximating distribution $q(\theta, \sigma^2, b_0 | \lambda)$ with a reasonable goodness of fit. From this point, Stochastic Variational Bayes will be used to update the posterior distribution as more data is observed to $p(\theta, \sigma^2, b_0 | y_{1:T+J}), J = 1, 2, \dots$ and provide the predictive density $p(y_{T+J+1} | y_{T+J})$.

\section{Discussion} \label{sec:Discussion}

Put main points and issues to investigate here

- When parameters change, models with non-simple filtering methods to obtain changes in states. ie variable parameter state space models.

\section{Timeline} \label{sec:Timeline}

Put this in!

\bibliographystyle{asa}
\bibliography{references}

\end{document}
\grid