### 2017-03-26 DLM/outline.tex
DLM additions

\documentclass[12pt,a4paper]{article}%
\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{url}
\usepackage{wasysym}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\setcounter{MaxMatrixCols}{30}
\providecommand{\U}[1]{\protect\rule{.1in}{.1in}}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\numberwithin{equation}{section}
\begin{document}

\section{Introduction}

\begin{itemize}
\item Compare MCMC methods with variational methods for forecasting a Dynamic Linear Model and a Stochastic Volatility Model. 
\item Use a range of different updating methods to adjust forecasts as new data is observed.
\item Compare compuation time and forecast accuracy for all approaches.
\end{itemize}

\section{Approaches Considered}

\textbf{MCMC Methods}
\begin{itemize}
\item Gibbs + Metropolis-Hastings MCMC.
\item Hamiltonian Monte Carlo.
\item Particle Filtering / QMLE / Extended Kalman Filter / Whatever else is used for SV.
\end{itemize}
\textbf{Updating for MCMC}
\begin{itemize}
\item Full updating - rerun MCMC after each new observation for best possible result.
\item No updating - just use multi step ahead forecasts via filtering.
\end{itemize}
\textbf{Variational Methods}
\begin{itemize}
\item Independent Gaussian - Transform parameters to real line eg. $\log(\sigma^2)$.
\item Full Rank Gaussian with real line transformation.
\item Sampling informed distributions - Copula model from Di{\ss}mann's algorithm.
\item Neural Network transform of $N(\boldsymbol{0}, I)$ posterior.
\item Stochastic Recurrent Neural Networks - SV Model only.
\item Alternative divergence measures eg. Expectation Propogation.
\end{itemize}
\textbf{Updating for VB}
\begin{itemize}
\item Update all parameters and latent variables.
\item Update global parameters and latent variables for new data - Hold previous latent variables fixed.
\item Do not update, just use filtering and multi step ahead forecasts.
\end{itemize}

\section{Simulation Results}

Simulate each estimation/update combination many times, calculate cumulative scores of new observations as well as compuation time for initial fit and each update.

\subsection{Dynamic Linear Model}

\begin{align}
y_t &= \mu + \beta x_t + \sigma_y \epsilon_t \\
x_t &= \gamma + \phi x_{t-1} + \sigma_x \nu_t
\end{align}

$\epsilon_t$ and $\nu_t \sim Z$

\subsection{Stochastic Volatility Model}

\begin{align}
y_t &= \exp\left(\alpha_t/2\right) \epsilon_t \\
\alpha_t &= \gamma + \phi \alpha_{t-1} + \sigma \nu_t
\end{align}

$\epsilon_t$ and $\nu_t \sim Z$.
Alternatively transform to another representaiton if that works better for filtering/QMLE etc.

\section{Empirical Application}
\begin{itemize}
\item DLM: Find something
\item SV: Intraday stock prices? 
\item Possibly have a multivariate $\textbf{y}_t$
\end{itemize}

\bibliographystyle{asa}
\bibliography{references}

\section{Discussion}

\end{document}