### 2017-09-21 Cars/basicModel.Rnw
Hamilton Filtering

\documentclass[12pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{alltt}%
\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{tikz}
%\usetikzlibrary{bayesnet}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\setcounter{MaxMatrixCols}{30}

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\setlength\parindent{0pt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\begin{document}

\section{The Basic Model}

Consider the motion model where x, y coordinates in eucliedean space can be determined by the history of coordinates as well as the current velocity, $v$, and steering angle, $\delta$, at time $t$ by
\begin{align}
x_t &= x_{t-1} + v_t \cos(\delta_t) \label{xEq} \\
y_t &= y_{t-1} + v_t \sin(\delta_t). \label{yEq}
\end{align}

The change in velocity equals acceleration, $a$, which is a stationary autoregressive process.
\begin{align}
v_t &= v_{t-1} + a_t \label{vEq} \\
a_t &= \phi a_{t-1} + \epsilon_t \label{aEq}
\end{align}
reducing to an ARIMA(1, 1, 0) process for velocity:
\begin{equation}
v_t - v_{t-1} = \phi(v_{t-1} - v_{t-2}) + \epsilon_t. \label{vARIMA}
\end{equation}

Finally, steering angle follows an AR(1) process with mean $\pi/2$.
\begin{equation}
\delta_t = \pi / 2 + \gamma (\delta_{t-1} - \pi / 2) + \eta_t \label{dEq}.
\end{equation}

Note that $v_t$ and $\delta_t$ in (\ref{xEq}) and (\ref{yEq}) are fully identifiable from observations of $\Delta x_t = x_t -  x_{t-1}$ and $\Delta y_t = y_t - y_{t-1}$,
\begin{align}
v_t &= v(\Delta x_t, \Delta y_t) = \sqrt{\Delta x_t^2 + \Delta y_t^2} \label{vEq2} \\
\delta_t &= \delta(\Delta x_t, \Delta y_t) = \tan^{-1}\left(\frac{\Delta y_t}{\Delta x_t} \right). \label{dEq2}
\end{align}

\iffalse
These can be substituted into (\ref{vARIMA}) and (\ref{dEq}) to give
\begin{align}
\sqrt{\Delta x_t^2 + \Delta y_t^2} &=  \sqrt{\Delta x_{t-1}^2 + \Delta y_{t-1}^2} + \phi \left(\sqrt{\Delta x_{t-1}^2 + \Delta y_{t-1}^2} - \sqrt{\Delta x_{t-2}^2 + \Delta y_{t-2}^2}\right) + \epsilon_t \label{vEq3} \\
\tan^{-1}\left(\frac{\Delta y_t}{\Delta x_t} \right) &= \pi / 2 + \gamma \left(\tan^{-1}\left(\frac{\Delta y_{t-1}}{\Delta x_{t-1}} \right) - \pi/2 \right) + \eta_t \label{dEq3}
\end{align}
\fi

Assuming that $\epsilon_t \sim N(0, \sigma^2_{\epsilon})$ and $\eta_t \sim N(0, \sigma^2_{\eta})$, using the notation that $x_{1:T} = \{x_t | t = 1, \dots T\}$, $y_{1:T} = \{y_t | t = 1, \dots T\}$, and $\theta = \{\gamma, \phi, \sigma^2_{\epsilon}, \sigma^2_{\eta}\}$, as well as treating $x_{1:3}$ and $y_{1:3}$ as given, the likelihood can be written as
\begin{equation}
p(x_{4:T}, y_{4:T} | \theta) = \prod_{t=4}^T \left(p_v(x_t, y_t | \theta) p_{\delta}(x_t, y_t | \theta) \right)
\end{equation}
where
\begin{align}
p_v(x_t, y_t | \theta) &\propto  \sigma_{\epsilon}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\epsilon}} \left(v(\Delta x_t, \Delta y_t) - (1+\phi) v(\Delta x_{t-1}, \Delta y_{t-1}) + \phi v(\Delta x_{t-2}, \Delta y_{t-2})\right)^2  \right\} \label{vDens} \\
p_{\delta}(x_t, y_t | \theta) &\propto  \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \pi/2 - \gamma \left(\delta(\Delta x_{t-1}, \Delta y_{t-1}) - \pi/2\right)\right)^2 \right\} \label{dDens} 
\end{align}
The following priors are used, where $IG$ denotes the inverse of a Gamma(shape, rate) distribution,
\begin{align*}
\gamma &\sim B(\alpha_{\gamma}, \beta_{\gamma}) \\
\phi &\sim B(\alpha_{\phi}, \beta_{\phi}) \\
\sigma^2_{\epsilon} &\sim IG(\alpha_{\epsilon}, \beta_{\epsilon}) \\
\sigma^2_{\eta} &\sim IG(\alpha_{\eta}, \beta_{\eta})
\end{align*}

A variational approximation is constructed for $\theta^* = \{\log(\gamma / (1-\gamma)), \log(\phi / (1-\phi)), \log(\sigma^2_{\epsilon}), \log(\sigma^2_{\eta})\}$, approximating the unknown posterior of $p(\theta^* | x_{1:T}, y_{1:T})$ with a multivariate normal distribution $q_{\lambda}(\theta^* | x_{1:T}, y_{1:T})$, choosing the set of paramaters $\lambda$ that minimises the Kullback-Leibler divergence from $q$ to $p$,
\begin{equation}
KL(q || p) = \int_{\theta^*} q_{\lambda}(\theta^* | x_{1:T}, y_{1:T}) \log \left( \frac{ q_{\lambda}(\theta^* | x_{1:T}, y_{1:T})}{p(\theta^* | x_{1:T}, y_{1:T})} \right) d_{\theta^*}.
\end{equation}

\begin{figure}[hb]
\centering
\includegraphics[width = 0.6\textwidth]{basicFit.png}
\end{figure}

\section{Adding a Markov Switching Variable}

The complexity of the steering process in (\ref{dEq}) is increased by introducing a two-state Markov switching variable $s_t$, where $s_t = 0$ indicates that the driver is continuing on their current path, and $s_t = 1$ will turn the car towards the centre of the lane. The new steering equation is described by
\begin{equation}
\delta_t = (1 - s_t) \delta_{t-1}  s_t (pi/2 + \zeta(x_{t-1} - M)) +  \eta_t \label{dEqMS}.
\end{equation}
where $M$ denotes the midpoint of the lane. Note that this now describes a random walk when $s_t = 0$, with large departures from $x_t = M$ controlled when $s_t = 1$.  $s_t$ is governed by the transition probabilities $\rho_{ij} = p(s_t = j | s_{t-1} = 1)$, with $\rho_{00} = 1 - \rho_{01}$ and $\rho_{10} = 1 - \rho_{11}$. As drivers are more likely to steer towards the centre of the lane the further away they are from the centre, $\rho_{10}$ and $\rho_{11}$ are given by 
\begin{equation}
\rho_{i1} = \frac{1}{1 + \exp\{-(\alpha_i + \beta_i (x_{t-1} - M)^2)\}} \mbox{ for } i = 0, 1.
\end{equation}
With this formulation, $s_t =  0$ represents a default driving state, and $s_t = 1$ represents the driver actively correcting for drift.

This formulation removes $\gamma$ from $\theta$, but adds five additional parameters, $\zeta, \alpha_0, \beta_0, \alpha_1$, and $\beta_1$, each with a $N(\mu_i, \tau_i^2)$ prior. The likelihood can be calculated by the Hamilton filter, and is given by
\begin{equation}
p(x_{4:T}, y_{4:T} | \theta) = \prod_{t=4}^T \left(\sum_{i=0}^1 \sum_{j=0}^1 \rho_{ij} \xi_{i, t-1} p_{v}(x_t, y_t | \theta) p_{\delta j}(x_t, y_t | \theta) \right)
\end{equation}
where
\begin{align}
p_{\delta 0}(x_t, y_t | \theta) &\propto  \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \delta(\Delta x_{t-1}, \Delta y_{t-1}) \right)^2 \right\} \label{dDensS0}, \\
p_{\delta 1}(x_t, y_t | \theta) &\propto   \sigma_{\eta}^{-1} \exp\left\{\frac{-1}{2 \sigma^2_{\eta}} \left(\delta(\Delta x_t, \Delta y_t) - \pi/2 - \zeta (x_{t-1} - M)\right)^2  \right\} \label{dDensS1},
\end{align}
and
\begin{align}
\xi_{j, t} &= P(s_t = j | x_0, y_0, \dots, x_{t}, y_{t}, \theta) \\
&= \frac{\sum_{i=0}^1\rho_{ij} \xi_{i, t-1} p_v(x_{t}, y_{t})p_{\delta j}(x_{t}, y_{t})}
{\sum_{i=0}^1 \sum_{j=0}^1 \rho_{ij} \xi_{i, t-1} p_{v}(x_t, y_t | \theta) p_{\delta j}(x_t, y_t | \theta)} \label{HF}.
\end{align}
The recursion in (\ref{HF}) is initialised at $\xi_3 = \{0.9, 0.1\}$.
\end{document}