### 2018-02-27 Cars/paper.tex
Paper Revision

  \documentclass[12pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{alltt}%
\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage[aboveskip=2pt]{subcaption}
\usepackage{amsmath}
\usepackage{amsthm}
%\usepackage{tikz}
%\usetikzlibrary{bayesnet}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\setcounter{MaxMatrixCols}{30}

\def\app#1#2{%
  \mathrel{%
    \setbox0=\hbox{$#1\sim$}%
    \setbox2=\hbox{%
      \rlap{\hbox{$#1\propto$}}%
      \lower1.1\ht0\box0%
    }%
    \raise0.25\ht2\box2%
  }%
}
\def\approxprop{\mathpalette\app\relax}

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\evensidemargin}{0.1in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.25in}
\setlength\parindent{0pt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\title{Online Updating of Variational Bayes for Heterogenous Forecasts of Vehicle Trajectory}

%TODO:
%Flesh out section 1
%Add authors, institutions, an abstract and bibtex references 
%Switch out manual references for Bibtex when everything else is done.

\begin{document}
\maketitle

\section{Introduction}
\label{sec:intro}

\begin{itemize}
\item Self driving vehicles are becoming more common, eg Waymo, Tesla, GM.
\item The naviation systems used to drive involve predicting the movement of the surrounding traffic to find the safest path for the self driving vehicle.
\item While self driving vehicles reduce collisions due to human error, safety can be further improved by better forecasts of the trajectory of nearby human controlled vehicles.
\item These models use deep neural networks (DNN) to extract the properties of surrounding vehicles from sensors (eg. Cameras and LiDAR), such as their position, velocity, and steering angle, see \citet{Woo2016a} and \citet{Tian2017}.
\item The self driving vehicle forecasts point estimates of the future trajectory of these surrounding vehicles based on the DNN results.
\item These can assume the surrounding vehicles will maintain their current angle and velocity \citet{Gindele2010}
\citet{Waymo2017}, \citet{Bautista2017}, but interest has developed in more advanced predictive models such as Neural Networks, Hidden Markov Models, and Support Vector Machines \citet{Woo2016b}, \citet{Ding2013}, \citet{Woo2017}, \citet{Geng2017}.
\item This paper takes a statistical approach, fitting time series models to two extracted variables, velocity and steering angle, of vehicles in the Next Generation Simulation (NGSIM) US Highway 101 Dataset.
\item It then uses Bayesian methods to forecast the distribution of the future velocity and steering angle, which can be transformed into a trajectory forecast.
\item A homogenous time series model shows a strong ability to predict the MAP trajectory
\item An important part of this modelling is to capture the uncertainty of car trajectories, as the self driving vehicle must react to the possibility of collision even when that possibility is relatively small.
\item There are many sources of heterogeneity in drivers, from their individual driving styles to the conditions of traffic directly around them.
\item Including heterogeneity in the model will allow the model to better adapt to drivers with more extreme behaviours, for example 'lead-footed' drivers can be assigned a larger variance on their acceleration while forecasts for more consistent drivers will benefit from a low variance.
\item A hierarchical model is employed to allow heterogeneity between different drivers, where each vehicle is given its own parameter vector with a shared prior distribution.
\item This model implies that any new vehicle observed by the self driving vehicle has its own parameter vector, related to previously observed vehicles through the shared prior.
\item The self driving vehicle must be able to estimate the parameter vector of newly observed cars, and update this parameter vector as it observes additional data.
\item Forecasts where the parameter vector is conditioned on this additional data should outperform forecasts conditioned on less data.
\item A three step strategy is employed to provide density forecasts for the future trajectory of newly observed vehicles:
\begin{enumerate}
\item Fit a hierarchical model with a large amount of vehicles to obtain the posterior distribution of each vehicles individual parameter vector and the shared distribution. This can occur before the self driving vehicle is on the road.
\item When a new vehicle is observed use the information in this shared distribution to model the associated new parameter vector. 
\item Update the parameter vector and all forecasts at short intervals.
\end{enumerate}
\item This paper uses standard Variational Bayes for the second step, and introduces an updating method for the third step; the posterior distribution of the new vehicles parameter vector is approximated, then this approximation is updated.
\item This only uses the data observed in-between updates, allowing the previously observed history of that vehicle to be discarded.
\item This is important for two reasons:
\begin{enumerate}
\item The model can be updated faster as less data must be processed, so forecasts can be made using as much recent data as possible.
\item Self driving vehicles take in up to 1GB of data a second (several websites report this as quotes from different people eg. Tech CEO, but I don't know if that's good enough for an academic citaiton), so discarding old data is useful.
\end{enumerate}
\item Show that a homogenous time series model has strong predictive performance compared to constant models / neural networks while also providing density forecasts.
\item Build a heterogenous version of the homogenous model, show that the density forecasts improve. 
\item Show that the Updating VB required to implement the heterogenous model in real time is comparable to MCMC and standard VB.
\end{itemize}

\section{Data processing}
\label{sec:dataProcessing}
Data is provided by the Next Generation Simulation (NGSIM) project, which recorded vehicles travelling along a 2235 feet section of the US 101 Freeway in Los Angeles, California from 7:50 am to 8:35 am on June 15th, 2005. Data was collected by seven static cameras and processed by Cambridge Systematics Inc. The result is coordinates relative to the start of the road section for 6101 vehicles collected every 100 milliseconds. The average number of observations per vehicle is 672, corresponding to a 67.2 seconds of data.
Figure \ref{fig:rawData} shows the paths of ten vehicles in the dataset, which travel along the five major lanes or the enter from the entry/exit lane to the right, with one vehicle changing from Lane 2 to Lane 1. Vehicles that entered or exited midway through the freeway section are excluded in this paper. There is a curvature to the road, with bends occuring between 500 and 1000 feet and 1800 and 2000 feet.

\begin{figure}
\centering
\includegraphics[width=0.65\textwidth]{carPath}
\caption{The path of ten vehicles in the dataset. The freeway is split into five lanes, and an additional entry/exit lane to the far right, through which two vehicles from this sample have entered. There is curvature to the road slightly distored by the aspect ratio, with bends occuring around 500 and 1800 feet.}
\label{fig:rawData}
\end{figure}

Modern vehicles are capable of tracking lane marking, and thus can compensate for the curvature of the road as they calculate the location of surrounding vehicles in coordinates relative to their own position (Thuy and Leon 2010), rather than relative to a static point as provided by NGSIM. To compensate for this, and so that the data is in a coordinate system similar to that of a self driving vehicle tracking surrounding cars, the coordinate system provided in the NGSIM dataset is transformed into a relative system where the road curvature is removed. This is facilitated by measuring the location of each vehicle relative to the estimate of the midpoint of each lane. \citet{Woo2016a} fit a polynomial curve to detected lane markings to build a local model of the road lane edges, and this idea is extended with the use of smoothing splines to build a model of the lane midpoints for the entire road section.

The midpoints for each of the five lanes are estimated separately, using 100 randomly sampled vehicles each that did not change their lane over the observed period. For each sampled vehicle $i$ and time since entering the road $t$, the total distance travelled is calcualted as 
\begin{equation}
\label{distance}
d_{i, t} = \sqrt{(x_{i, t}) - x_{i, 1})^2 + (y_{i, t} - y_{i, 1})^2}
\end{equation}
From this distance metric, the average $x$ and $y$ coordinates at each given lane and for a given distance are each estimated by a pair of smoothing splines calculated with the R `stats' package. The smoothing splines are functions of the distance travelled and estimate the lane midpoints as  $\hat{x}_{i, t} = f(d_{i, t})$ and $\hat{y}_{i, t} = g(d_{i, t})$. Each remaining vehicle in the dataset uses the lane midpoint estimates $\{\hat{x}^*_{i, t}, \hat{y}^*_{i, t}\}$ fit from the spline model assocaiated with its starting lane to calculate relative coordinates $\{x^*_{i, t}, y^*_{i, t}\}$, where $y^*_{i, t}$ denotes the distance travelled along the road, and $x^*_{i, t}$ denotes the devitation from the lane midpoint, by
\begin{align}
x^*_{i, t} &= \mbox{sign}\left(\tan^{-1}\left(\frac{g'(d_{i, t}) }{f'(d_{i, t})}\right) - \tan^{-1}\left(\frac{\hat{y}_{i, t} - y_{i, t}}{\hat{x}_{i, t} - x_{i, t}} \right)\right)\sqrt{(x_{i, t}-\hat{x}_{i, t})^2 + (y_{i, t} - \hat{y}_{i, t})^2)} \label{xRel} \\
y^*_{i, t} &= d_{i, t}. \label{yRel}
\end{align}
The vehicles used to estimate the five spline models are excluded for any other purpose.
\begin{figure}
\centering
\includegraphics[width = 0.5\textwidth]{relCoord}
\caption{An example of the equation in (\ref{xRel}). The estimated trajectory of the lane midpoint is given by the solid curve, while a vehicle $i$ at a given point in time $t$ is given by the blue dot at $\{x_{i, t}, y_{i, t}\}$. This vehicle has travelled a distance equivalent to the red dot at $\{\hat{x}_{i, t}, \hat{y}_{i, t}\}$. The gradient of the midpoint at this point is given by the dashed line, which intercepts the x-axis at an angle of $\lambda = \tan^{-1}\left(\frac{g'(d_{i, t}) }{f'(d_{i, t})}\right)$. The dotted line from the blue dot to the red intercepts the x-axis at an angle of $\psi = \tan^{-1}\left(\frac{\hat{y}_{i, t} - y_{i, t}}{\hat{x}_{i, t} - x_{i, t}} \right)$. The sign of the angle $\lambda - \psi$ determines whether the vehicle is to the left or right side of the lane, and thus has a negative or positive relative coordinate $x^*_{i, t}$. The absolute value of $x^*_{i, t}$ is the distance between the two dots.}
\label{fig:relCoord}
\end{figure}
Once relative coordinates are extracted, the relative position of any vehicle $i$ at any time $t$ is determined by its initial position $\{x^*_{i, 1}, y^*_{i, 1}\}$, as well as the history of that driver's inputs: velocity $\{v_{i, s} | s = 2, \dots, t\}$ and steering angle, $\{\delta_{i, s} | s = 2, \dots, t\}$ by the trigonometric relationship in Figure \ref{fig:motion}:
\begin{align}
x^*_{i, t} &= x^*_{i, t-1} + v_{i, t} \cos(\delta_{i, t}) \label{xEq} \\
y^*_{i, t} &= y^*_{i, t-1} + v_{i, t} \sin(\delta_{i, t}) \label{yEq},
\end{align}
where $\delta_{i, t} = \pi/2$ denotes that the car has no lateral relative movement and $x^*_{i, t} = x^*_{i, t-1}$.
From this relationship, the driver's inputs are be extracted by
\begin{align}
\delta_{i, t} &= \tan^{-1}\left(\frac{(y^*_{i, t} - y^*_{i, t-1})}{(x^*_{i, t} - x^*_{i, t-1})} \right) \label{dEq} \\
v_{i, t} &= \sqrt{(x^*_{i, t} - x^*_{i, t-1})^2 + (y^*_{i, t} - y^*_{i, t-1})^2} \label{vEq}.
\end{align}
It is then useful to extract the acceleration, $a_{i, t}$ by the simple relation
\begin{equation}
\label{aEq}
a_{i, t} = v_{i, t} - v_{i, t-1}. 
\end{equation}
Data for vehicle $i$ up to time $T$ is defined as $z_{i, 1:T} = \{x^*_{i, s}, y^*_{i, s}, | s = 1, \dots, T\}$ and data from multiple drivers are then collected as $z_{1:N} = \{z_{i, 1:T_i} | i = 1, \dots, N\}$.
\begin{figure}
\centering
\includegraphics[width = 0.4\textwidth]{motion}
\caption{A vehicle at $x_{t-1}, y_{t-1}$ at time $t-1$ will be at $x_t, y_t$ at time $t$ if its angle over this period is $\delta_t$ and velocity is $v_t$.}
\label{fig:motion}
\end{figure}

\section{Trajectory Forecasts}
\label{sec:trajForecasts}

A crucial step in the safe operation of self driving vehicles is to forecast the future trajectory of surrounding vehicles so that collisions can be avoided. To produce these forecasts the data is split to place vehicles into two sets: a training set used to infer the posterior distribution of a parameter set before the self driving vehicle is on the road; and a test set used for forecasting the positions of vehicles observed while driving. The parameters associated with the vehicles in the test set may be inferred, but this inference is time constrained. If inference takes $M$ seconds, a forecast after observing a vehicle for $T$ seconds can only include data up to time $T - M$ to infer the posterior distribution of any parameters.

The remainder of Section \ref{sec:trajForecasts} will introduce the models used, while Sections \ref{sec:Inference} and \ref{sec:UVB} will build an inferential framework to facilitate inference of any additional parameters in constant time as $T$ increases to apply to online trajectory forecasting.

\subsection{Homogenous Time Series Model}
\label{subsec:homogenous}

The framework provided by (\ref{xEq}) and (\ref{yEq}) allows forecasts of $\{x^*_{t}, y^*_{t} | t = T + 1, \dots, T+H\}$ to be obtained given the position and velocity of a vehicle at time $T$, $\{x^*_T, y^*_T, v_T\}$, and forecasts of the future values of 
$\{a_{t}, \delta_{t} | t = T + 1, \dots, T+H\}$.

As data is observed every 100 milliseconds, large changes in $a_{i, t}$ and $\delta_{i, t}$ can occur over multiple time periods. This can be observed in Figure \ref{fig:dynamics}, where the left panel plots the mean value of
\begin{equation}
A_{i, t} = \frac{a_{i, \tau_i + t}}{a_{i, \tau_i}}
\label{amax}
\end{equation}
where $\tau_i = \arg \underset{t}{\max}|a_{i, t}|$. On average, a vehicle will increase in acceleration for the 400 milliseconds before reaching maximum acceleration. The right panel plots the mean value of
\begin{equation}
D_{i, t} = \frac{\delta_{i, \omega_i + t} - \pi/2}{\delta_{i, \omega_i} - \pi/2}
\label{dmax}
\end{equation}
where $\omega_i = \arg \underset{t}{\max}|\delta_{i, t} - \pi/2|$, which similarly demonstrates that large absolute deviations in steering angle from $pi/2$ are associated with smaller deviations in the same direction in the the previous 300 milliseconds.

\begin{figure}[h]
\centering
\includegraphics[width = 0.95\textwidth]{dynamics}
\caption{Left: Mean values of acceleration as a percent of a vehicles maximum absolute acceleration for the 1000 milliseconds before, and 500 milliseconds after, a vehicle reaches its maximum absolute acceleration. Right: The analogous plot for the steering angle, with $\delta_t - \pi/2$ as a percentage of a vehicles maximum absolute value of $\delta_t - \pi/2$. The few observations before a variable reaches its maximum often have increasing values for the same variable.}
\label{fig:dynamics}
\end{figure}

These dynamics imply that changes in acceleration and angle are predictable given the recent history of behaviour, and thus auto-regressive processes of orders $p$ and $q$ for $a_t$ and $\delta_t$ are employed to model this, with
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{j} a_{i, t-j} + \sigma_{\epsilon} \epsilon_{i, t} \label{aAR} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta} \eta_{i, t} \label{dAR}
\end{align}
for each $i = 1, \dots, N$, where $\epsilon_{i, t}$ and $\eta_{i, t}$ are independently and identically distributed according to a standard normal distribution.

The parameter vector is transformed to $\mathbb{R}^{p + q + 2}$ and collected as 
\begin{equation*}
\label{thetaVec}
\theta = \{\phi_{1}, \dots, \phi_{p}, \gamma_{1}, \dots, \gamma_{q}, \log(\sigma^{2}_{\epsilon}), \log(\sigma^{2}_{\eta})\},
\end{equation*}
where
\begin{equation}
\label{indPrior}
p(\theta) \sim \mathcal{N}\left(\mu, \Sigma \right)
\end{equation}
where $\mu$ is a vector with an entry of $0$ for each $\phi$ and $\gamma$, and an entry of $-5$ for each log-variance, as the scale of changes in the data is very small. $\Sigma = 10 \mathbb{I}$, where $\mathbb{I}$ is the identity matrix. This is referred to as the homogenous model throughout the paper.

\subsection{Heterogenous Time Series Models}
\label{subsec:heterogenous}

The behaviour of a driver, and the dynamics of their actions, may depend on many individual features such as their experience, type of vehicle, personality, and the characteristics of traffic surrounding them. Heterogeneity is allowed by replacing the shared $\theta$ parameter vector with $N$ individual parameter vectors $\theta_i$ so that
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{i, j} a_{i, t-j} + \sigma_{\epsilon, i} \epsilon_{i, t} \label{aAR2} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{i, j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta, i} \eta_{i, t}. \label{dAR2}
\end{align}
Using the prior in (\ref{indPrior}) for each $\theta_i$ formulates what is referred to as the independent model where each vehicle posterior is independent of all other vehicles, so $p(\theta_i | z_{1:N}, \theta_{j \neq i}) = p(\theta_i | z_i)$. An alternative approach is to augment the prior distribution of each $\theta_i$ with parameters $\beta$ in a hierarchical model, and infer the distribution of the $p(\theta_{1:N}, \beta | z_{1:N})$. The top panel of Figure \ref{fig:HierSingleKDE} plots a kernel density estimate of the posterior means of each $\theta_i, i = 1, \dots, 2000$ using the independent model with $p = q = 2$ at time $T = 500$. There is a wide range of values for the posterior mean of each variable, indicating heterogeneity across the drivers in the sample. The density of posterior means for both variance parameters, $\sigma^2_{\epsilon}$ and $\sigma^2_{\eta}$, exhibit strong positive skewness, while the density of posterior means of both $\phi_1$ and $\phi_2$ are multimodal. Due these features, a $K$ component mixture of multivariate normal distributions is chosen for $p(\theta_i | \beta)$, augmenting each driver with an auxiliary variable $k_i = 1, \dots, K$ such that 
\begin{equation}
\label{mixPrior}
\theta_i | k_i = j \sim N(\mu_j, \Sigma_j),
\end{equation}
and
\begin{equation}
k_i \sim \mbox{Multinomial}\left(\pi_1, \dots, \pi_{K}\right)
\end{equation}
for each $i = 1, \dots, N$. Furthurmore, for each $j = 1, \dots, K$,  
\begin{align}
\mu_j &\sim \mathcal{N}\left(\bar{\mu}_j, \Omega_j\right), \\
\Sigma_j &\sim \mbox{Inverse Wishart}\left(\mbox{Degrees of Freedom } \tau_j, \mbox{Scale } \Psi_j\right), \\
\boldsymbol{\pi} &\sim \mbox{Dirichlet}\left(\alpha_1 = \alpha_2 = \dots = \alpha_K\right)
\end{align}
The elements of $\beta$ are $\{\mu_j, \Sigma_j, \pi_j | j = 1, \dots, K\}$. Each $\bar{\mu}_j$ contains a $0$ for each $\phi$ and $\gamma$, and $-5$ for each log-variance, each $\Omega_j = 10 \mathbb{I}$, each $\tau_j = 6$, each $\Psi_j = \mathbb{I}$ and finally each $\alpha_j = 1$. The bottom panel of Figure \ref{fig:HierSingleKDE} plots the marginal distributions for $p(\theta_{2001} | \beta, z_{1:2000})$ for the hierarchical model with $p = q = 2$ and $K = 6$ using the same vehicles as the independent model in the top panel. The marginal distributions capture the skewness and multimodality present in the densities of posterior means from the independent model.

\begin{figure}[ht]
\centering
\includegraphics[width = 0.95\textwidth]{HierSingleKDE.png}
\caption{Top: A Kernel Density Estimate applied to the posterior means for each $\theta_i | z_i, i = 1, \dots, 2000$ where $p = q = 2$ from the independent model, sampled by Metropolis-Hasting MCMC. Bottom: $p(\theta_i | \beta, z_{1:2000})$ for the hierarchical model with $K = 6$ using the same vehicles sampled by Metropolis-Hasting MCMC. Each individual component in the mixture is denoted by the coloured densities. The hierarchical model has captured the multimodality and skewness present in the independent posterior means.}
\label{fig:HierSingleKDE}
\end{figure}

\subsection{Encountering Additional Vehicles}
\label{subsec:additionalVehicles}

For the remainder of this paper $z_{1:N}$ and $\theta_{1:N}$ refers to the data collected from, and associated parameter vectors for, the $N$ vehicles that are available before the self driving vehicle is on the road, while the subscript $i$ is reserved for some $i > N$ and refer to the data and parameter vector for any additional vehicle encountered while the self driving vehicle is on the road.  Inference about $\theta_{1:N}$, the individual parameter vectors for the independent and hierarchical models, as well as for $\theta$, the shared homogenous parameter vector and $\beta$, the hierarchical hyperparameters, conditional on $z_{1:N}$ are assumed to be available with an unlimited computational budget.

A significant difference in the application of the time series models is through the relationship between the posterior distribution formed from observations of the first $N$ vehicles and the parameter vector for the additional vehicle, $\theta_{i}$. Under the homogenous model $\theta$ is shared across all vehicles and hence $\theta_{i} = \theta$. In this case $p(\theta | z_{1:N}, z_{i, 1:T})$ will not be significantly different to $p(\theta | z_{1:N})$ due to the high precision of the posterior distribution for a sufficiently large $N$. Figure (\ref{fig:homogPosterior}) plots the posterior marginal distributions from $N = 2000$ vehicles highlighting the posterior precision.

\begin{figure}
\centering
\includegraphics[width = 0.95\textwidth]{homogPosterior}
\caption{The converged Metropolis Hastings MCMC samples of the marginal posterior distributions $p(\theta | z_{1:N})$ for the homogenous model with $N = 2000$ and $p = q = 2$. Each variable has very high posterior precision.}
\label{fig:homogPosterior}
\end{figure}

The independent model implies that the posterior distribution for $\theta_i$ is independent of $\theta_{1:N}$ and $z_{1:N}$ by
\begin{equation}
p(\theta_{i}| z_{1:N}, z_{i,1:T}, \theta_{1:N}) = p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_i) p(\theta_i).
\label{indepNewCar}
\end{equation}
The posterior for $\theta_i$ using the hierarchical model, assuming conditional independence between each $\theta_j$ for $j = 1, \dots, N$ and $\theta_i$, given $\beta$, $z_{1:N}$, and $z_i$, is
\begin{equation}
\label{hierNewCar}
p(\theta_{i} | z_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) \int_{\beta} p(\theta_{i} | \beta) p (\beta | z_{1:N}) d\beta.
\end{equation}
As N is large the sampled posterior $p(\beta | z_{1:N})$ has high precision, and $\int_{\beta} p(\theta_{i} | \beta) p (\beta | z_{1:N}) d\beta$ can be replaced with $p(\theta_{i} | \hat{\beta})$ in (\ref{hierNewCar}) where $\hat{\beta}$ is a point estimate such as the mean or maximum of $p(\beta | z_{1:N})$. This approximation results in the closed form expression
\begin{equation}
\label{hierNewCar2}
p(\theta_{i} | z_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) p(\theta_{i} | \hat{\beta}).
\end{equation}
(\ref{indepNewCar}) and (\ref{hierNewCar2}) can be combined with an inference procedure to allow the parameter set for any vehicle observed while driving to be inferred, and used in the forecast equation (\ref{predictive}) to produce online forecasts of the trajectory of those vehicles that allows individual heterogeneity to be included.

\section{Bayesian Inference}
\label{sec:Inference}

Given a series of data observed up to time $T$, $z_{1:T}$, the Bayesian forecast distribution associated with some future time $T+h$ is characterized by the conditional density
\begin{equation}
\label{predictive}
p(z_{T+h} | z_{1:T}) =\int p(z_{T+h}|z_{1:T}, \theta) p(\theta | z_{1:T}) d\theta.
\end{equation}
To obtain this distribution, the posterior density for $\theta$, given by
\begin{equation}
\label{posterior}
 p(\theta | z_{1:T}) = \frac{p(z_{1:T}|\theta)p(\theta)}{\int p(z_{1:T}|\theta)p(\theta) d\theta}
\end{equation}
must first be inferred. Generally the analytical solution to (\ref{posterior}), and hence to (\ref{predictive}), will be unavailable due to the complex functional form of the distributions involved.

In this section two alternative methods for computing the desired posterior distribution will be reviewed: Markov Chain Monte Carlo (MCMC) and Variational Bayes (VB). In brief, MCMC is used to create a sample from $p(\theta | z_{1:T})$, with any function of $\theta$ that is desired estimated from that sample. In contrast, VB replaces $p(\theta | z_{1:T})$ with a parametric approximation, denoted by $q_{\lambda}(\theta | z_{1:T})$, where $\lambda$ is a vector of auxiliary parameters associated with the approximation that may depend on the observations $y_{1:T}$.

\subsection{Markov Chain Monte Carlo}
\label{subsec:MCMC}

There are many types of MCMC algorithms, with arguably the simplest and most commonly used one being the Gibbs sampler. The Gibbs sampler algorithm iteratively samples the components of the $k-$dimensional parameter vector $\theta$ via each of the so-called full conditional distributions as follows,
\begin{align}
&p(\theta_1 | \theta_2, \dots, \theta_k, z) \nonumber \\
&p(\theta_2 | \theta_1, \theta_3, \dots, \theta_k, z) \nonumber \\
&\vdots \nonumber \\
&p(\theta_k | \theta_1, \dots, \theta_{k-1}, z). \nonumber
\end{align}
Under mild regularity conditions (see, e.g., \citet{Tierney1994}) and with enough iterations of the Markov chain that results from the Gibbs sampler, these samples converge in distribution to $p(\theta | z)$. Samples taken before the MCMC converges to the posterior must be discarded, and the remaining samples may have strong dependence between consecutive draws of the same parameter due to the Markov nature of the algorithm. The computation time for each iteration and the overall number of iterations required to accruately summarise the posterior distribution is problem specific and typically increases with the number of parameters in the model. The full conditional distributions cannot be recognised for each of the time series models proposed in Section \ref{sec:trajForecasts} and a Metropolis-Hastings-within-Gibbs (MH) step is utilised instead \citep{Gilks1995}.

In MH MCMC, a candidate $\theta_j^{(c)}$, where $\theta_j$ may be any scalar or vector subset of $\theta$, is drawn from a proposal distribution $r(\theta_j)$ and accepted by the sampler with probability
\begin{equation}
\min \left\{ 1, \frac{p(\theta_j^{(c)} | \theta_{i \neq j}^{(a)}, z)}{p(\theta_j^{(a)} | \theta_{i \neq j}^{(a)}, z)} \times \frac{r(\theta_j^{(a)} | \theta_j^{(c)})}{r(\theta_j^{(c)} | \theta_j^{(a)})} \right\},
\label{MHaccept}
\end{equation}
where the superscript $(a)$ denotes the most recently accepted value of the corresponding subset of $\theta$. If the sampler rejects a candidate value the previous value $\theta_j^{(a)}$ is repeated. Each iteration of Metropolis-Hastings-within-Gibbs MCMC includes one candidate draw for each element of $\theta$. Choice of proposal distribution is left to the user with arguably the most simple being the Normal Random Walk proposal,
\begin{equation}
\theta_j^{(c)} \sim N(\theta_j^{(a)}, \Sigma_j)
\label{RWprop}
\end{equation}
as $r(\theta_j^{(c)} | \theta_j^{(a)}) = r(\theta_j^{(a)} | \theta_j^{(c)})$. Performance of Random Walk Metropolis Hastings MCMC (RWMH-MCMC) then depends on the acceptance rate of the proposal distribution, largely influenced by the value of $\Sigma_j$. \citet{Garthwaite2016} propose an adaptive algorithm that increases or decreases $\Sigma_j$ depending if the most recent candidate was rejected or accepted respectively so that a target acceptance rate is asymptotically approached as the number of MCMC iterations approaches infinity.

\subsection{Variational Bayes}
\label{subsec:VB}

A typically faster, albeit approximate, alternative to MCMC based inference is VB \citep{Jordan1999}. VB posits a family of parametric approximating distributions $q_{\lambda}(\theta | z)$, parameterised by an auxiliary vector $\lambda$, that share the same support as the true posterior distribution $p(\theta | z)$. Note that the family $q_{\lambda}(\theta | z)$ does not depend on $z$, but this notation is used to make it clear that this distribution is an approximation for $p(\theta | z)$. A member of approximating family is chosen to minimise some error function, typically the Kullback-Leibler (KL) divergence from $q_{\lambda}(\theta | z)$ to $p(\theta | z)$ \citep{Kullback1951}, given by $KL[q_{\lambda}(\theta | z)\hspace{.1cm}||\hspace{.1cm}p(\theta | z_{1:T})]$. The KL divergence is defined by
\begin{equation}
\label{KL-def}
KL[q_{\lambda}(\theta | z)\hspace{.1cm}||\hspace{.1cm}p(\theta | z_{1:T})] = E_{q_{\lambda}(\theta | z)} \left[ \log(q_{\lambda}(\theta | z)) - \log(p(\theta | z)) \right],
\end{equation}
and is a non-negative, asymmetric measure of the discrepancy between $p(\theta | z)$ and $q_{\lambda}(\theta | z)$  that will be equal to zero if and only if $p(\theta | z) = q_{\lambda}(\theta | z)$ almost everywhere \citep{Bishop2006}.
Typically (\ref{KL-def}) cannot be evaluated, and Monte-Carlo estimates as
\begin{equation}
\label{KL-MC}
KL[q_{\lambda}(\theta | z)\hspace{.1cm}||\hspace{.1cm}p(\theta | z_{1:T})] \approx \frac{1}{M}\sum_{i=1}^M \left(\log(q_{\lambda}(\theta_i |z)) - \log(p(\theta | z)) \right)
\end{equation}
where $\theta_i \sim q_{\lambda}(\theta |z))$ are compuationally infeasible due to the inclusion of the term $p(\theta | z)$. Instead VB uses the Evidence Lower Bound (ELBO), denoted by $\mathcal{L}(q, \lambda)$, as an error function where
\begin{equation}
\label{ELBO}
\mathcal{L}(q, \lambda) = E_{q_{\lambda}(\theta |z)} \left[\log(p(\theta, z)) - \log(q_{\lambda}(\theta |z))\right],
\end{equation}
which is evaluated with Monte-Carlo estimates
\begin{equation}
\label{ELBO-MC}
\mathcal{L}(q, \lambda) \approx \frac{1}{M} \sum_{i=1}^M \left(\log(p(\theta_i, z)) - \log(q_{\lambda}(\theta_i |z)) \right)
\end{equation}
where $\theta_i \sim q_{\lambda}(\theta |z))$. The ELBO is equal to the negative KL divergence plus a constant, and hence maximising (\ref{ELBO}) with respect to $q_{\lambda}(\theta | z)$ is equivalent to minimising (\ref{KL-def}).

\subsection{Stochastic Gradient Ascent}
\label{subsec:SGA}
For exponential family likelihood models, $p$, with a factorisable approximation, $q$, the characteristics of the surface of the ELBO can be exploited for optimisation in Mean Field Variational Bayes \citep{Ghahramani2000, Wainwright2008}, but for more general distributions the surface of the ELBO and its stochastic estimate are unknown. Maximisation proceeds by optimising only the auxiliary parameters $\lambda$ for a fixed distribution family $q$ with stochastic gradient ascent (SGA). SGA repeatedly takes Monte-Carlo estimates of the gradient of the ELBO with respect to $\lambda$, $\delta\mathcal{L}(q, \lambda) / \delta \lambda$, as $\widehat{\delta\mathcal{L}(q, \lambda) / \delta \lambda}$ and applies updates of the form
\begin{equation}
\label{gradientAscent}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}} \bigg\rvert_{\lambda = \lambda^{(m)}}
\end{equation}
until the change from $\mathcal{L}(q, \lambda^{(m)})$ to $\mathcal{L}(q, \lambda^{(m+1)})$ falls below some pre-specified threshold \citep{Hoffman2013}. Intuitively, individual elements of $\lambda$ will increase if the estimate of the slope of $\mathcal{L}(q, \lambda^{(m)})$ is positive at the current point, and will decrease if that estimate is negative, until each element of $\lambda$ reaches a point where the slope is zero. This procedure is guaranteed to converge to a local maximum \citep{Robbins1951} if the sequence $\rho^{(m)}, m = 1, \dots, \infty$ satisfies
\begin{align}
&\sum_{m=1}^{\infty} \rho^{(m)} =  \infty \\
&\sum_{m=1}^{\infty} (\rho^{(m)})^2 <  \infty.
\end{align}
In this paper the sequence $\rho^{(m)}$ is provided by the Adam algorithm of \citet{Kingma2015b}.

There are two choices for the Monte Carlo estimator of $\delta\mathcal{L}(q, \lambda) / \delta \lambda$, the score estimator of \citet{Ranganath2014}, 
\begin{equation}
\label{scoreDeriv}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{SC} = \sum_{j = 1}^M \frac{\delta \log(q_{\lambda}(\theta_j | z))}{\delta \lambda} \left(\log(p(\theta_j, z)) - \log(q_{\lambda}(\theta_j | z)) \right),
\end{equation}
where $\theta_j \sim q_{\lambda}(\theta | z)$ and the reparameteterised estimator of \citet{Kingma2014}. Reparameterisation rephrases Variational Bayes as a search for the parameters $\lambda$ of some differentiable function $f(\cdot, \lambda)$ that minimises the Kullback Leibler divergence from some distribution $r(\epsilon)$ with zero free parameters to a transformation of the posterior distribution, $p_{\epsilon}(\theta = f(\epsilon, \lambda) | z)$. Examples of $f$ and $r(\epsilon)$ include treating $\theta$ as location scale transformation from a standard normal $\epsilon$, or an inverse-CDF transformation from a uniform$(0, 1)$ $\epsilon$. The distribution $p(\theta, z)$ is written as
\begin{equation}
\label{rpDist}
p(\theta, z) = p(f(\epsilon, \lambda), z)|J(f(\epsilon, \lambda))|
\end{equation}
where $J(f(\epsilon, \lambda))$ is the Jacobian Matrix of the transformation $f(\epsilon, \lambda)$, and the reparameterised ELBO can be found by substituting (\ref{rpDist}) into (\ref{ELBO}),
\begin{equation}
\label{rpELBO}
\mathcal{L}(q, \lambda) = E_{r(\epsilon)} \bigg[\log(p(f(\epsilon,\lambda), z)|J(f(\epsilon, \lambda))|) - \log(r(\epsilon))\bigg].
\end{equation}
The gradient of the reparameterised ELBO with respect to $\lambda$ of 
\begin{align}
\label{rpELBODeriv}
\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda} &= \frac{\delta}{\delta \lambda} \bigg( E_{r(\epsilon)} \bigg[\log\big(p(f(\epsilon,\lambda), z)|J(f(\epsilon, \lambda))|\big) - \log(r(\epsilon))\bigg] \bigg) \nonumber \\
&= E_{r(\epsilon)} \left[ \frac{\delta}{\delta \lambda} \bigg(\log(p(f(\epsilon,\lambda), z)) + \log(|J(f(\epsilon, \lambda))|) - \log(r(\epsilon)) \bigg)\right] \nonumber \\
&= E_{r(\epsilon)} \left[ \frac{\delta \log(p(f(\epsilon,\lambda), z))}{\delta f(\epsilon,\lambda)} \frac{\delta f(\epsilon,\lambda)}{\delta \lambda}  + \frac{\delta \log(|J(f(\epsilon, \lambda))|)}{\delta \lambda} \right].
\end{align}
This form leads to the reparameterised estimator of $\delta\mathcal{L}(q, \lambda) / \delta \lambda$
\begin{equation}
\label{rpDeriv}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{RP} = \sum_{j = 1}^M \frac{\delta f(\lambda, \epsilon_j)}{\delta \lambda} \frac{\delta \log(p(\theta, z))}{\delta \theta} \bigg\rvert_{\theta = f(\lambda, \epsilon_j)} + \frac{\delta J(\lambda, \epsilon_j)}{\delta \lambda}, 
\end{equation}
where $\epsilon_j \sim r(\epsilon)$.

These Monte Carlo estimators converge to the true derivative at a rate $O(\sqrt{M})$, so choosing the number of samples per estimate, $M$, involves a trade-off between the computation time per gradient estimate and the stochastic noise present in each estimate. An increased value of $M$ will reduce the estimator variance, and generally reduce the number of iterations required for the ELBO to converge, at a linear increase in computation time per iteration. To reduce the variance of the Monte Carlo estimator, following \citet{Gunawan2017}, Randomised Quasi Monte Carlo (RQMC) is used, where numbers in the unit hypercube are generated according to a Sobol Sequence \citep{Sobol1967} which are then randomised using the scrambled net method \citep{Matousek1998}, and transformed using the inverse-CDF of $q_{\lambda}(\theta | z)$ or $r(\epsilon)$. Using RQMC allows the gradient estimators to converge to the true derivative at the increased rate of $O(M)$.

\section{Updating Variational Bayes}
\label{sec:UVB}

The self driving vehicle is constantly observing data about the movements of the surrounding vehicles. After observing vehicle $i$ until time $S$, $z_{i, 1:S}$ has been observed and can be incorporated into the Variational Bayes posterior approximation $q_{\lambda_S}(\theta_{i} | z_{i, 1:S})$, suppressing the conditional dependence on $z_{1:N}$. At time $T > S$ an additional $T - S$ data points from vehicle $i$ have been observed, and forecasts could be improved by incorporating these to form the posterior approximation, $q_{\lambda_T}(\theta_{i} | z_{i, 1:T})$, where the $S$ and $T$ subscripts on $\lambda$ are included to differentiate the auxiliary parameter vector conditioned on data up to times $S$ and $T$. To facilitate this posterior update, an Updating Variational Bayes (UVB) mechanism is introduced where only $z_{i, S+1:T}$ needs to be processed.

Given $p(\theta_{i} | z_{i, 1:S})$ from (\ref{posterior}), the posterior distribution at time $T$, $p(\theta_{i} | z_{i, 1:T})$ is given by Bayes rule as
\begin{equation}
\label{updatePost}
p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S})
\end{equation}
Standard methods to evaluate the posterior, such as Variational Bayes or Markov Chain Monte Carlo, require the evaluation of the right hand side of (\ref{updatePost}), however computation of $p(\theta_{i} | z_{i, 1:S})$ is often infeasible. Without being able to evalute the right hand side, posterior inference requires evaluation of the prior distribution and full sample $z_{i, 1:T}$. To avoid this, Variational Bayesian Updating replaces $p(\theta_{i} | z_{i, 1:S})$ with the analytical approximation $q_{\lambda_S}(\theta_{i} | z_{i, 1:S})$. The update (\ref{updatePost}) is replaced with
\begin{equation}
\label{updateEqApprox}
p(\theta_{i} |  z_{i, 1:T}) \approxprop p(z_{i, S+1:T} | \theta_{i})q_{\lambda_S}(\theta_{i} | z_{i, 1:S})
\end{equation}
The ELBO gradient estimators to construct the updated Variational Bayes approximation at time $T$,  $q_{\lambda_T}(\theta_{i} | z_{i, 1:T})$, can be obtained by substituting (\ref{updateEqApprox}) into the score estimator (\ref{scoreDeriv}) or the reparameterised estimator (\ref{rpDeriv}) of $\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda} $. The updating score estimator is given by
\begin{align}
\widehat{\frac{\delta\mathcal{L}(q, \lambda_T)}{\delta \lambda_T}}_{SCU} &= \sum_{j = 1}^M \frac{\delta \log(q_{\lambda_T}(\theta_{i, j} | z_{i, 1:T}))}{\delta \lambda_T} \nonumber \\
&\times \left(\log(q_{\lambda_S}(\theta_{i, j} | z_{i, 1:S}) - \log(q_{\lambda_T}(\theta_{i} z_{i, 1:T})) \right) \label{scoreUpdate}
\end{align}
where $\theta_{i, j} \sim q(\theta_{i} | z_{i, 1:T})$. Similarly, the updating reparameterised estimator is given by
\begin{equation}
\label{rpUpdate}
\widehat{\frac{\delta\mathcal{L}(q, \lambda_T)}{\delta \lambda_T}}_{RPU} = \sum_{j = 1}^M \frac{\delta f(\lambda_T, \epsilon_j)}{\delta \lambda_T} \frac{\delta \log(q_{\lambda_S}(\theta_{i} |z_{i, 1:S}))}{\delta \theta_{i}} \bigg\rvert_{\theta_{i} = f(\lambda_T, \epsilon_j)} + \frac{\delta J(\lambda_T, \epsilon_j)}{\delta \lambda_T},
\end{equation}
where $\epsilon_j \sim r(\epsilon)$. If this process is repeated periodically, $S$ will grow at the same rate as $T$, and updating VB is $O(1)$ rather than $O(T)$ as required by re-using the full sample without the approximation step in (\ref{updateEqApprox}).

\section{Forecast Evaluation}
\label{sec:eval}

After obtaining the posterior distributions for each model conditioned on 2000 vehicles, with $p = q = 2$ and $K = 6$, forecasts are created and evaluated on a further 1373 vehicles. For each of these vehicles $i > 2000$, at each point in time $T = 100, 110, \dots, 450$ and forecast horizon $h = 1, \dots, 30$, point estimates of $z_{i, T+h}$ are obtained from every naive model, while the time-series models are used to provide forecast densities of $p(z_{i, T+h} | z_{1:2000}, z_{i, 1:T})$ and point estimates from the MAP of this density. As data is observed every 100 milliseconds, this equates to forecasting the next three seconds of movement after every second of observing the vehicle. The forecast density for the homogenous model is created using the MCMC samples of $p(\theta | z_{1:2000}, z_{i, T})$, while the independent and hierarchical models produce $p(\theta_{i} | z_{1:2000}, z_{i, 1:T})$ through three methods:
\begin{enumerate}
\item Posterior sampling with RWMH-MCMC,
\item VB fitting $q_{\lambda_T}(\theta_{i} | z_{1:2000}, z_{i, 1:T})$ to the complete $z_{i, 1:T}$ using the original prior distribution,
\item UVB fitting $q_{\lambda_T}(\theta_{i} | z_{1:2000}, z_{i, 1:T})$ using $q_{\lambda_S}(\theta_{i} | z_{1:2000}, z_{i, 1:S})$ and data $z_{i, S+1:T}$ as described in Section \ref{sec:UVB}.
\end{enumerate}
A summary of these approaches is provided in Table \ref{tableAlg}. Point estimate forecasts are evaluated by their Euclidean Error,
\begin{equation}
\mbox{EE}_{i, T, h} = \sqrt{\left(\hat{x}^*_{i, T+h} - x^*_{i, T+h} \right)^2 + \left(\hat{y}^*_{i, T+h} - y^*_{i, T+h} \right)^2},
\label{eucError}
\end{equation}
where $\{\hat{x}^*_{i, T+h}, \hat{y}^*_{i, T+h}\}$ is a forecast of $\{x^*_{i, T+h}, y^*_{i, T+h}\}$, while density forecasts are evalutated by their logscore,
\begin{equation}
\mbox{LS}_{i, T, h} = \log \left(p\left(x^*_{i, T+h}, y^*_{i, T+h} | z_{1:2000}, z_{i, 1:T} \right) \right)
\label{logscore}
\end{equation}

\begin{center}
\begin{table}[h]
\resizebox{\textwidth}{!}{%
\begin{tabular}{| l | c | c |}
\hline
& Independent Model & Hierarchical Model \\
\hline
MCMC & \multicolumn{2}{c|}{RWMH-MCMC jointly drawing the entire $\theta$ vector using \citet{Garthwaite2016} to control the} \\
& \multicolumn{2}{c|}{variance matrix of the Multivariate Gaussian proposal distribution to obtain a 23.4\% acceptance rate for $\theta$ draws.} \\
\hline
VB  & Variational Bayes using (\ref{rpDeriv}) to update parameters & Variational Bayes using (\ref{scoreDeriv}) to update parameters\\
Standard &  of a Multivariate Gaussian approximation with &  of an approximation formed as a six component mixture\\
&non-zero covariance using $M = 25$, where $f$ is a &  of diagonal variance Multivariate Gaussians using $M = 50$. \\
&location scale transform from a standard normal $r(\epsilon)$. &\\
\hline
VB  & Variational Bayes using (\ref{rpUpdate}) to update parameters & Variational Bayes using (\ref{scoreUpdate}) to update parameters\\
Standard &  of a Multivariate Gaussian approximation with &  of an approximation formed as a six component mixture\\
&non-zero covariance using $M = 25$, where $f$ is a&  of diagonal variance Multivariate Gaussians using $M = 50$. \\
&location scale transform from a standard normal $r(\epsilon)$. &\\
\hline
\end{tabular}}
\caption{Details of the algorithms used to produce posterior distributions for each method and prior combination}
\label{tableAlg}
\end{table}
\end{center}

\subsection{Naive Forecast Models}
\label{subsec:Naive}

Nine naive forecast models are provided where forecasts of $a_t$ and $\delta_t$ are constants equal to $\hat{a}$ and $\hat{\delta}$ for $t = T+1, \dots, T+H$. The classification of each naive model is provided in Table \ref{tableNaive} as a combination of one estimator for $\hat{a}$ and one estimator for $\hat{\delta}$ at time $T$. Acceleration estimators are the most recent acceleration, $\hat{a} = a_T$, the average acceleration over the most recent second, $\hat{a} = 0.1 \sum_{s = T-9}^T a_s$, and zero acceleration $\hat{a} = 0$, equivalent to constant velocity. Similarly, steering angle estimators are the most recent angle, $\hat{\delta} = \delta_T$, the average angle over the most recent second, $\hat{\delta} = 0.1 \sum_{s = T - 9}^T \delta_s$, and $\hat{\delta} = \pi/2$, corresponding to the vehicle to driving directly forward.
\begin{table}
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
& $\hat{\delta}_t = \delta_T$ & $\hat{\delta}_t = 1/10 \sum_{s=T-9}^T \delta_s$ & $\hat{\delta}_t = \pi/2$ \\
\hline
$\hat{a}_t = a_T$ & Naive 1 & Naive 2 & Naive 3 \\
$\hat{a}_t = 1/10\sum_{s=T-9}^T a_s$ & Naive 4 & Naive 5 & Naive 6\\
$\hat{a}_t = 0$ & Naive 7 & Naive 8 & Naive 9 \\
\hline
\end{tabular}
\end{center}
\caption{Classification of naive forecast models by future velocity and steering angle estimators.}
\label{tableNaive}
\end{table}

\subsection{Recurrent Neural Networks}
\label{subsec:RNN}
A Recurrent Neural Network (RNN) is constructed with inputs following \citet{Ding2013}, the ten most recent lags of vehicle position, angle, velocity, and time headway to the preceding vehicle. The RNN outputs the next thirty changes in the vehicle position, is trained with the Adam optimiser and a a loss function $EE_{i, T, 30}$ given by (\ref{eucError}). As of writing this report forecast results from this model are not competitive with the naive models and are not discussed further.

\subsection{Results}
\label{subsec:Results}

\begin{itemize}
\item Discuss the updating method relative to standard VB, change in logscore in Figure \ref{fig:updateCost}
\item Plot the average convergence time as a function of $T - S = 1, 5, 10, 20$ and $T = 100, 110, \dots,  450$. Running now for Figure \ref{fig:timing}
\end{itemize}

Mean Eucliedean Error for each model at forecast horizons of one, two, and three seconds ($h = 10, 20$ and $30$) are provided in Figure \ref{fig:PredError}. Naive model performance is split into three groups according to the choice of acceleration forecast, and for a given $\hat{a}_t$ the choice of $\hat{\delta}_t$ has not substantially changed the forecast error. Zero acceleration naive models (Naive 7, 8, and 9) perform the best, followed by the averaged acceleration models (Naive 4, 5, and 6) and finally the current acceleration models (Naive 1, 2, and 3). At all forecast horizons the error for each implementation of each of the time series models are significantly lower than every naive model.

Figure (\ref{fig:PredErrorZ}) truncates Figure (\ref{fig:PredError}) to include only the time series models, and with the exception of the independent model with UVB inference, each time series model and inference implementation produces a forecast with a similar amount of error. This implies that there is no systematic increase or decrease in MAP forecast accuracy for choosing either the homogenous or introducing heterogeneity with the heirarchical model. As the independent model is the heterogenous model that provides the least accurate forecasts it is not discussed further.

\begin{figure}[ht]
\centering
\includegraphics[width = 0.8\textwidth]{predictiveError.png}
\caption{Mean predictive Euclidean error in metres for $p(z_{i, T+h} | z_{i, 1:T})$ for $h \in \{10, 20, 30\}$ and $T \in \{100, 110, \dots, 450\}$ for each model. Each time series model is significantly more accurate than the naive constant acceleration / angle models, with little discernable difference between the time series models.}
\label{fig:PredError}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width = 0.8\textwidth]{predictiveErrorZoom.png}
\caption{The same results as Figure \ref{fig:PredError} with the Naive models excluded. Most of the time series model have very similar MAP forecast accuracy, with the UVB / Independent time series combination performing slightly weaker}
\label{fig:PredErrorZ}
\end{figure}

There are systematic differences in the level of certainty in forecasts provided by the heirarchical and homogenous models as measured by their logscores. The key differentiation between the models is the heirarchical models ability to assign each vehicle individual variance parameters,  $\sigma^2_{\epsilon, i}$ for $a_t$, and $\sigma^2_{\eta, i}$ for $\delta$. Figure (\ref{fig:posVarMean}) plots the range of posterior means for these parameters from with RWMH-MCMC sampling with $T = 450$, with the homogenous posterior means indicated with vertical red lines. There is a large amount of positive skewness in the distribution of variances, which may have caused the homogenous model to estimate a variance that is too large for most individual cars. The homogenous variance posterior means are greater than $80\%$ and $91.5\%$ of heirarchical variance posterior means for acceleration and the steering angle respectivley. This large variance has little impact on the location of the MAP of $p(z_{i, T+h} | z_{1:N}, z_{i, 1:T})$, but impacts the level of uncertainty in the forecast density. Figure \ref{fig:varSplit} shows a detailed breakdown of the median increase in predictive logscore obtained from using the hierarchical model with updating VB inference over the homogenous model at the three second forecast horizon. Vehicles are split into bivariate quintiles according to their posterior variance means for acceleration along the x-axis and angle along the y-axis. The hierarchical model trajectory forecasts for drivers in the lower variance quntiles benefit the most from including heterogeneity. However the median difference is zero, or even negative, for forecasts of drivers in the fifth quintiles. This paper argues that heterogenous modelling is still useful in this scenario, as the individual vehicle variance estimates allow identification of vehicles with high variance, and potentially more erratic and dangerous driving styles.

\begin{figure}[h]
\centering
\includegraphics[width = 0.95\textwidth]{posVarMean}
\caption{Kernel Density Estimates for the posterior mean of the variance parameters, $\sigma^2_{\eta, i}$ and $\sigma^2_{\epsilon, i}$ for each of the 1387 forecasted cars fit by Metropolis-Hastings MCMC at $T = 450$ using the hierarchical model, compared to the homogenous posterior mean in red. Small amounts of cars with high variance has skewed the homogenous estimates to be too large for the majority of the vehicles forecasted.}
\label{fig:posVarMean}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width = 0.95\textwidth, height = 0.4\textheight]{varSplit}
\caption{Difference in logscore between the hierarchical model with posterior inference using UVB and the homogenous model. Results are split by variance quintiles for acceleration (x-axis) and angle (y-axis). There are substantial improvements in logscore, and thus forecast certainty, for vehicles with a low variance in either variable. Vehicles in the fifth quintiles, close to the homogenous variance estimates, benefit instead from the higher precision associated with the homogenous posterior distribution due to the large $N$.}
\label{fig:varSplit}
\end{figure}

The benefits of heterogenous modelling is only useful if the posterior distribution $p(\theta_{i} | z_{1:N}, z_{1:T})$, or its variational approximation $q_{\lambda_T}(\theta_{i} | z_{1:N}, z_{1:T})$ can be inferred in the short time frames demanded by self driving vehicles. Figure (\ref{fig:timing}) demonstrates the average computation time required for standard VB (Black) and UVB (coloured), where UVB is first fit to $100$ observations then repeatedly updated every $T-S$ observations for different values of $T-S$. The compuation time for standard VB increases linearly with $T$, while UVB is constant; posterior inference avabale at time $T$ utilising UVB can be conditioned on more data than standard VB. 
Parallelising the $M$ Monte-Carlo samples per iteration can further reduce computation time and make UVB feasible for more complex models. Figure (\ref{fig:updateCost}) demonstrates that implementing UVB over standard VB does not lead to a decrease in forecast logscores, plotting the difference between the heirarchical forecast logscores obtained by UVB and standard VB for several ranges of $T$. Augmenting self driving vehicles with hierarchical time series models to forecast the trajectory of surrounding traffic, and repeatedly updating the posterior distribution and forecasts, can be feasibly implemented with UVB.
\begin{figure}[htp]
\centering
\includegraphics[width = 0.75\textwidth]{timing}
\caption{Average time to converge for Standard VB (Black) and UVB (Coloured) with four different update lengths given by $T - S$, truncated to 25 seconds. Each UVB procedure originally fits the model to the first 100 data points and updates this every $T - S$ data points. VB algorithms are ran with $M = 50$ on one CPU core. Convergence time could be reduced through parallelisation.}
\label{fig:timing}
\end{figure}
\begin{figure}[htp]
\centering
\includegraphics[width = 0.75\textwidth, height = 0.3\textheight]{updateCost}
\caption{Differences in predictive logscores for $p(z_{i, T+h} | z_{i, 1:T})$ for each $h = 1, \dots, 30$ and $T = 100, 110, \dots, 450$ between updating and standard Variational Bayes. The differences are typically small, and UVB can improve the logscore.}
\label{fig:updateCost}
\end{figure}

\newpage

\section{Conclusion}
\begin{itemize}
\item The use of auto-regressive models for forecasting future acceleration and angles drastically improves forecasts of the future trajectory compared to the competing models.
\item There is evidence of heterogeneity between vehicles, and allowing this with the heirarchical model allows behaviour specific to an individual driver to be inferred, improving logscores and thus reducing forecast uncertainty. 
\item Heterogeneity particularly benefits forecasts for low variance drivers, but also enables high variance drivers to be identified.
\item Online forecasting with the heterogeneous model requires constant model fits and updates, computation of which can be simplified without a substantial performance hit by UVB.
\item UVB is more broadly applicable to any kind of online inference for time series data.
\end{itemize}

\newpage
\bibliographystyle{asa}
\bibliography{references}

\end{document}