### 2018-03-08 Cars/midcanPresentation.tex
presentation

\documentclass[8pt]{beamer}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}

\usepackage{alltt}
\usepackage{natbib}
\usepackage{array}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage[]{algorithm2e}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\usepackage{bbm}
\usepackage{tikz}
\usepackage{amssymb}
\usepackage{bm}
\setbeamertemplate{caption}[numbered]

\usetheme{Warsaw}
\defbeamertemplate*{footline}{shadow theme}
{%
  \leavevmode%
  \hbox{\begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm plus1fil,rightskip=.3cm]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertframenumber\,/\,\inserttotalframenumber\hfill\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.5\paperwidth,ht=2.5ex,dp=1.125ex,leftskip=.3cm,rightskip=.3cm plus1fil]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle%
  \end{beamercolorbox}}%
  \vskip0pt%
}

\setbeamertemplate{headline}{%
\leavevmode%
  \hbox{%
    \begin{beamercolorbox}[wd=\paperwidth,ht=2.5ex,dp=1.125ex]{palette quaternary}%
    \insertsectionnavigationhorizontal{\paperwidth}{\hskip0pt plus1filll}{\hskip0pt plus1filll}
    \end{beamercolorbox}%
  }
}

\newcounter{saveenumi}
\newcommand{\seti}{\setcounter{saveenumi}{\value{enumi}}}
\newcommand{\conti}{\setcounter{enumi}{\value{saveenumi}}}
\resetcounteronoverlays{saveenumi}

\title[Online Updating of Variational Bayes]{Online Updating of Variational Bayes for Heterogeneous Forecasts of Vehicle Trajectory}
\author[Nathaniel Tomasetti]{Nathaniel Tomasetti}
\date{ }
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\begin{frame}
\titlepage
\centering
Supervised by Catherine Forbes and Anastasios Panagiotelis
\end{frame}

\begin{frame}
\tableofcontents
\end{frame}

\begin{frame}
\section{Framework}
\frametitle{Thesis Framework}
Focus: Two sets of data
\begin{itemize}
\item Stuff Here about the general themes
\item Use of exact inference to improve VB
\item Hierarchical models with exact inference
\item Some new unit with time constrained inference
\item Want to use MCMC info to improve VB
\end{itemize}
\end{frame}

\begin{frame}
\section{Inference}
\frametitle{Bayesian Inference}
Note: reframe discussion with two sets of data from previous slide
\begin{itemize}
\item The posterior distribution for parameters $\theta$, $p(\theta | \textbf{z})$, can be inferred with Markov Chain Monte Carlo (MCMC). MCMC:
\begin{enumerate}
\item Allows 'exact' inference as the number of iterations increases
\item Can be very slow
\end{enumerate}
\item We focus on settings where MCMC inference for the additional unit is too slow to be useful.
\item Instead, Variational Bayes (VB) is used. VB:
\begin{enumerate}
\item Approximates the posterior with an analytical alternative $q_{\lambda}(\theta | \textbf{z})$.
\item Chooses the distribution $q$ that minimises a divergence funciton
\item Is typically much faster than MCMC.
\end{enumerate}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Kullback Leibler Divergence and Stochastic Optimisation}
\begin{itemize}
\item The typical divergence is the Kullback Leibler Divergence from $q$ to $p$.
\end{itemize}
\begin{equation}
\label{KL-def}
KL[q_{\lambda}(\theta | \textbf{z})\hspace{.1cm}||\hspace{.1cm}p(\theta | \textbf{z})] = E_{q_{\lambda}(\theta | \textbf{z})} \left[ \log(q_{\lambda}(\theta | \textbf{z})) - \log(p(\theta | \textbf{z})) \right],
\end{equation}
\begin{itemize}
\item Monte-Carlo simulations are usually unavailable due to the $p(\theta | \textbf{z})$ term.
\end{itemize}
\begin{equation}
\label{ELBO}
\mathcal{L}(q, \lambda) = E_{q_{\lambda}(\theta | \textbf{z})} \left[\log(p(\theta, \textbf{z})) - \log(q_{\lambda}(\theta | \textbf{z}))\right],
\end{equation}
\begin{itemize}
\item Stochastic Estimates are available through Monte-Carlo:
\end{itemize}
\begin{equation}
\label{ELBO-MC}
\mathcal{L}(q, \lambda) \approx \frac{1}{M} \sum_{j=1}^M \left(\log(p(\theta_{j}, \textbf{z})) - \log(q_{\lambda}(\theta_{j} | \textbf{z})) \right)
\end{equation}
where $\theta_{j} \sim q_{\lambda}(\theta | \textbf{z})$.
\end{frame}

\begin{frame}
\frametitle{Gradient Estimators}
\begin{itemize}
\item For most $p$ and $q$ only Stochastic Gradient Ascent is available for optimisation. Repeatedly apply:
\begin{equation}
\label{gradientAscent}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}} \bigg\rvert_{\lambda = \lambda^{(m)}}
\end{equation}
\item ELBO gradients can be estimated by the Score Estimator:
\begin{equation}
\label{scoreDeriv}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{SC} = \sum_{j = 1}^M \frac{\delta \log(q_{\lambda}(\theta_{j, i} | \textbf{z}))}{\delta \lambda} \left(\log(p(\theta_{j, i}, \textbf{z})) - \log(q_{\lambda}(\theta_{j, i} | \textbf{z})) \right),
\end{equation}
where $\theta_j \sim q_{\lambda}(\theta | \epsilon)$.
\item Or the Reparametersied Estimator, where $\theta = f(\epsilon, \lambda)$:
\begin{equation}
\label{rpDeriv}
\widehat{\frac{\delta\mathcal{L}(q, \lambda)}{\delta \lambda}}_{RP} = \sum_{j = 1}^M \frac{\delta \theta}{\delta \lambda} \frac{\delta \log(p(\theta, \textbf{z}))}{\delta \theta} \bigg\rvert_{\theta = f(\lambda, \epsilon_j)} + \frac{\delta J(\lambda, \epsilon_j)}{\delta \lambda}, 
\end{equation}
where $\epsilon_j \sim q(\epsilon)$.
\end{itemize}
\end{frame}

\begin{frame}
\section{Online Heterogenous Forecasting: Problem}
\subsection{Motivation}
\frametitle{Self-Driving Vehicles}
\begin{itemize}
\item Self-Driving Vehicles are a thing soon
\item The navigation system has three steps
\begin{enumerate}
\item Detect surrounding traffic
\item Forecast surrounding traffic
\item Navigate through safely
\end{enumerate}
\item Forecasting has not had much work
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Aims of the first paper}
\begin{itemize}
\item What we plan on doing with this first paper
\item Overall goals and stuff
\item Motivate heterogeneity
\item Note heterogeneity needs inference on vehicles as they are encountered
\item There is not much time to do that
\item Develop an approximation to the Bayesian Update Equation so we can infer parameters for new vehicles online
\item Also mention something on density forecasting would be nice
\end{itemize}
\end{frame}

\begin{frame}
\subsection{Data}
\frametitle{Next Generation Simulation}
\begin{itemize}
\item This is a dataset created for macro traffic modelling
\item It records all of the cars all of the time
\item The road is curved
\item Before and After of road straightening project
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Forecasting in the framework}
\begin{itemize}
\item That is a terrible title.
\item Split the data into two sets: train/test
\item Cars observed beforehand, cars encountered while driving
\end{itemize}
\begin{figure}
\centering
\includegraphics[width = 0.95\textwidth]{problem}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Trigonometric Motion Equations}
\begin{align}
x^*_{i, t} &= x^*_{i, t-1} + v_{i, t} \cos(\delta_{i, t}) \label{xEq}, \\
y^*_{i, t} &= y^*_{i, t-1} + v_{i, t} \sin(\delta_{i, t}) \label{yEq}.
\end{align}
Picture
\end{frame}

\begin{frame}
\frametitle{Extracting Driver Inputs}
\begin{align}
\delta_{i, t} &= 
     \begin{cases}
       \tan^{-1}\left(\frac{(y^*_{i, t} - y^*_{i, t-1})}{(x^*_{i, t} - x^*_{i, t-1})} \right)  &\quad\text{if }x^*_{i, t} \neq x^*_{i, t-1} \\
       \frac{\pi}{2} &\quad\text{if } y^*_{i, t} > y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       -\frac{\pi}{2} &\quad\text{if } y^*_{i, t} < y^*_{i, t-1} \mbox{ and } x^*_{i, t} = x^*_{i, t-1} \\
       \delta_{i, t-1} &\quad\text{otherwise,} \\ 
     \end{cases} \label{dEq} \\
v_{i, t} &= \sqrt{(x^*_{i, t} - x^*_{i, t-1})^2 + (y^*_{i, t} - y^*_{i, t-1})^2} \label{vEq}.
\end{align}
\begin{equation}
\label{aEq}
a_{i, t} = v_{i, t} - v_{i, t-1}. 
\end{equation}
\begin{itemize}
\item Why other variables are useful
\item Acceleration is good because relative velocity / non-stationarity
\end{itemize}
\end{frame}

\begin{frame}
\subsection{Modelling}
\frametitle{Auto-regressive Models for Trajectory Forecasting}
\begin{itemize}
\item Evidence of dynamics behaviour
\item Modelling approachs / relationship to test train set
\item Motivate the use of the three different models
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{A Homogeneous Approach}
Talk about what is going on here
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{j} a_{i, t-j} + \sigma_{\epsilon} \epsilon_{i, t} \label{aAR} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta} \eta_{i, t} \label{dAR}
\end{align}
\begin{equation*}
\label{thetaVec}
\theta = \{\phi_{1}, \dots, \phi_{p}, \gamma_{1}, \dots, \gamma_{q}, \log(\sigma^{2}_{\epsilon}), \log(\sigma^{2}_{\eta})\}.
\end{equation*}
\begin{equation}
\label{indPrior}
\theta \sim \mathcal{N}\left(\mu, \Sigma \right)
\end{equation}
\begin{itemize}
\item The Posterior
\item Why are we doing this
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The Independent Heterogeneity Model}
Maybe different drivers behave differently?
\begin{align}
a_{i, t} &= \sum_{j = 1}^p \phi_{i, j} a_{i, t-j} + \sigma_{\epsilon, i} \epsilon_{i, t} \label{aAR2} \\
\delta_{i, t} &= \pi/2 + \sum_{j = 1}^q \gamma_{i, j} (\delta_{i, t-j} - \pi/2) + \sigma_{\eta, i} \eta_{i, t}. \label{dAR2}
\end{align}
This breaks down into a bunch of independent posteriors
\begin{equation}
p(\theta_{1:N} \mid z_{1:N}) = \prod_{i=1}^N p(\theta_{i} \mid z_{i}),
\end{equation}
The training set isn't terribly useful.
\begin{equation}
p(\theta_{i}| \textbf{z}_{1:N}, z_{i,1:T}, \theta_{1:N}) = p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_i) p(\theta_i).
\label{indepNewCar}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
Hierarchical Models share information 
\begin{itemize}
\item Introduce Auxiliary Variable K
\item Sharing of the information
\end{itemize}
\begin{equation}
\label{mixPrior}
\theta_i | k_i = j \sim N(\mu_j, \Sigma_j),
\end{equation}
and
\begin{equation}
k_i \mid \beta \sim \mbox{Multinomial}\left(\pi_1, \dots, \pi_{K}\right)
\end{equation}
\begin{align}
\mu_j &\sim \mathcal{N}\left(\bar{\mu}_j, \Omega_j\right), \\
\Sigma_j &\sim \mbox{Inverse Wishart}\left(\mbox{Degrees of Freedom } \tau_j, \mbox{Scale } \Psi_j\right), \\
\boldsymbol{\pi} &\sim \mbox{Dirichlet}\left(\alpha_1 = \alpha_2 = \dots = \alpha_K\right).
\end{align}
\end{frame}

\begin{frame}
\frametitle{The Clustered Heterogeneity Model}
The new vehicle depends on the old vehicles through $\beta$.
\begin{equation}
\label{hierNewCar}
p(\theta_{i} | \textbf{z}_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) \int_{\beta} p(\theta_{i} | \beta) p (\beta | \textbf{z}_{1:N}) d\beta.
\end{equation}
The $\beta$ posterior is precise for large $N$
\begin{equation}
\label{hierNewCar2}
p(\theta_{i} | \textbf{z}_{1:N}, z_{i, 1:T}) \propto p(z_{i, 1:T} | \theta_{i}) p(\theta_{i} | \hat{\beta}).
\end{equation}
\end{frame}

\begin{frame}
\begin{itemize}
\item That picture showing the IH / CH fits
\item Evidence of heterogeneity and the suitability of the mixture model
\end{itemize}
\end{frame}

\begin{frame}
\section{Updating Variational Bayes}
\frametitle{Updating Variational Bayes}
\begin{itemize}
\item Motivation
\begin{figure}[ht]
\centering
\includegraphics[width = 0.95\textwidth]{timeUpdate}
\end{figure}
\item Notation: $q_{\lambda_S}(\theta_{i} | z_{i, 1:S})$ and $q_{\lambda_T}(\theta_{i} | z_{i, 1:T})$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Approximating Bayesian Updating}
\begin{equation}
\label{updatePost}
p(\theta_{i} | z_{i, 1:T}) \propto p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S})
\end{equation}
\begin{itemize}
\item This only works if the right hand side can be evaluated
\item VB Gradient Estimators requires the joint distribution
\begin{equation}
\label{updateJoint}
p(\theta_{i}, z_{i, 1:T}) = p(z_{i, S+1:T} | \theta_{i})p(\theta_{i} | z_{i, 1:S})
\end{equation}
\item We will replace it with this
\begin{equation}
\label{ApproxJoint}
\hat{p}(\theta_{i},  z_{i, 1:T}) = p(z_{i, S+1:T} | \theta_{i})q_{\lambda_S}(\theta_{i} | z_{i, 1:S}).
\end{equation}
\item Sources of error: How close $q$ is to $p$
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Updating Gradient Estimators}
The updating score estimator:
\begin{align}
\widehat{\frac{\delta\mathcal{L}(q, \lambda_T)}{\delta \lambda_T}}_{USC} &= \sum_{j = 1}^M \frac{\delta \log(q_{\lambda_T}(\theta_{i, j} | z_{i, 1:T}))}{\delta \lambda_T} \nonumber \\
&\times \left(\log(q_{\lambda_S}(\theta_{i, j} | z_{i, 1:S}) - \log(q_{\lambda_T}(\theta_{i, j} | z_{i, 1:T})) \right) \label{scoreUpdate}
\end{align}
where $\theta_{i, j} \sim q(\theta_{i} | z_{i, 1:T})$. 
The updating reparameterised estimator:
\begin{equation}
\label{rpUpdate}
\widehat{\frac{\delta\mathcal{L}(q, \lambda_T)}{\delta \lambda_T}}_{URP} = \sum_{j = 1}^M \frac{\delta f(\lambda_T, \epsilon_j)}{\delta \lambda_T} \frac{\delta \log(q_{\lambda_S}(\theta_{i} |z_{i, 1:S}))}{\delta \theta_{i}} \bigg\rvert_{\theta_{i} = f(\lambda_T, \epsilon_j)} + \frac{\delta J(\lambda_T, \epsilon_j)}{\delta \lambda_T},
\end{equation}
where $\epsilon_j \sim q(\epsilon)$.
\end{frame}

\begin{frame}
\section{Online Heterogeneous Forecasting: Results}
\frametitle{Point Forecasts}
\begin{itemize}
\item Introduce Naive / RNN models?
\item Some stuff about point forecasts: Both pictures?
\item Note that, on average, the homogenous model is as good as others
\item Some stuff about different inference frameworks
\item MAPs are not the whole story
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Heterogeneous Variance Distributions}
\begin{itemize}
\item Some stuff about variance
\item A note on erratic driver detection
\item Two different pictures to go in here
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Updating Variational Bayes Results}
\begin{itemize}
\item Compare to regular VB
\item Time Complexity
\item Difference in logscores
\end{itemize}
\end{frame}

\begin{frame}
\section{Extensions}
\frametitle{Extending the Framework}
\begin{itemize}
\item Copula Project
\item Other Applications
\end{itemize}
\end{frame}

\end{document}