### 2018-08-28 betel.tex
betel

  \documentclass[12pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb
\hbadness=99999

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{alltt}%
\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)

%2multibyte Version: 5.50.0.2960 CodePage: 1252
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[centertags]{amsmath}
\usepackage{graphicx}%
\usepackage{natbib}
\usepackage{color}
\usepackage[dvipsnames,svgnames*]{xcolor}
\usepackage{array}
\usepackage[hidelinks]{hyperref}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{amsthm}
\usepackage[]{algorithm2e}
%\usepackage{tikz}
%\usetikzlibrary{bayesnet}
\usepackage{url}
\usepackage{ulem}
\usepackage{afterpage}
\setcounter{MaxMatrixCols}{30}

\def\app#1#2{%
  \mathrel{%
    \setbox0=\hbox{$#1\sim$}%
    \setbox2=\hbox{%
      \rlap{\hbox{$#1\propto$}}%
      \lower1.1\ht0\box0%
    }%i
    \raise0.25\ht2\box2%
  }%
}
\def\approxprop{\mathpalette\app\relax}

\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.08in}
\setlength{\evensidemargin}{0.08in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength\parindent{0pt}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

\title{Variational Bayesian Exponentially Tilted Empirical Likelihood}
\begin{document}
\maketitle

\section{Background}

\citet{Chib2017} introduces Bayesian inference based on moment conditions, basing inference on the satisfication of the $d-$dimensional moment conditions,

\begin{equation}
\label{emp:moments}
E \left[ g(Y, \theta) \right] = 0,
\end{equation}
where the expectation is with respect to the true data generation process $p$.

The likelihood is replaced with the exponentially tilted empirical likelihood (ETEL), defined by
\begin{equation}
\label{emp:etel}
p(y_{1:T} | \theta) = \prod_{t=1}^T p^*_t(\theta)
\end{equation}
where each $p^*_t$ is chosesn to minimise the KL divergence between $(p_1, p_2, \ldots, p_T)$ and the empirical probabilities $(1/T, 1/T, \ldots 1/T)$ subject to the conditions
\begin{equation}
\label{emp:sumRestrict}
\sum_{t=1}^T p_t = 1
\end{equation}
and
\begin{equation}
\label{emp:momentRestrict}
\sum_{t=1}^T p_t g(y_t, \theta) = 0.
\end{equation}
It can be shown that the solution to this is given by
\begin{equation}
\label{emp:solnP}
p^*_t(\theta) = \frac{\exp \{\hat{\lambda}^{\prime} g(y_t, \theta) \}}{ \sum_{s=1}^T \exp \{\hat{\lambda}^{\prime} g(y_s, \theta) \}}
\end{equation}
where
\begin{equation}
\label{emp:solnLambda}
\hat{\lambda} = \lambda(\theta) = \arg \underset{\lambda}{\min} \frac{1}{T} \sum_{t=1}^T \exp \{\lambda^{\prime} g(y_t, \theta) \}.
\end{equation}

This results in the ETEL posterior
\begin{equation}
\label{emp:post}
p(\theta | y_{1:T}) \propto p(\theta) \prod_{t=1}^T\frac{\exp \{\hat{\lambda}^{\prime} g(y_t, \theta) \}}{ \sum_{s=1}^T \exp \{\hat{\lambda}^{\prime} g(y_s, \theta) \}}.
\end{equation}

\citet{Chib2017} proposes a Metropolis Hastings sampler for the posterior, following the steps
\begin{enumerate}
\item Draw $\theta$ from a suitable candidate distribution
\item Solve (\ref{emp:solnLambda}) for this draw
\item Calculate (\ref{emp:post}) as part of the Metropolis Hastings acceptance probability
\item Accept or reject $\theta$ as per standard MH
\end{enumerate}
This sounds quite slow, and may benefit from VB.
\\

\section{Implementation of VB}

Reparameterised and Stein VB requires the derivative of $\log(p(\theta, y_{1:T})$ with respect to $\theta$ as part of the ELBO gradient estiamtor. This derivative is complicated by $\hat{\lambda}$ being a function of $\theta$.
\\

If, for every $\theta \in \Theta$, the solution to  (\ref{emp:solnLambda}) is unique, and, defining $h(\theta, \lambda(\theta)) =  \frac{1}{T} \sum_{t=1}^T \exp \{\lambda^{\prime} g(y_t, \theta) \}$, the hessian matrix $\frac{\partial^2  h(\theta, \lambda(\theta))}{\partial \lambda^2}$ is positive-defiite, the optimal $\hat{\lambda}$ is found by satisfying the first order condition
\begin{equation}
\label{emp:lambdaDeriv}
\frac{\partial h(\theta, \lambda(\theta))}{\partial \lambda} = 0.
\end{equation}
Taking the derivative of both sides of this with respect to $\theta$ results in
\begin{equation}
\label{emp:thetaDeriv}
\frac{\partial^2  h(\theta, \lambda(\theta))}{\partial \lambda \partial \theta^{\prime}} + \frac{\partial^2  h(\theta, \lambda(\theta))}{\partial \lambda \partial \lambda^{\prime}} \frac{\partial \lambda(\theta)}{\partial \theta} = 0,
\end{equation}
where
\begin{equation}
\label{emp:hessian}
\frac{\partial^2  h(\theta, \lambda(\theta))}{\partial \lambda \partial \lambda^{\prime}} = \frac{1}{T} \sum_{t=1}^T g(y_t, \theta) g(y_t, \theta)^{\prime} \exp\{\hat{\lambda}^{\prime} g(y_t, \theta) \}
\end{equation}
and
\begin{equation}
\label{emp:crossDeriv}
\frac{\partial^ 2  h(\theta, \lambda(\theta))}{\partial \lambda \partial \theta^{\prime}} = \frac{1}{T} \sum_{t=1}^T \exp\{\hat{\lambda}^{\prime} g(y_t, \theta) \} \left( \mathbb{I} + g(y_t, \theta) \hat{\lambda}^{\prime} \right) \frac{\partial g(y_t, \theta)}{\partial \theta}.
\end{equation}
(\ref{emp:thetaDeriv}) may be solved for $\partial \lambda(\theta) / \partial \theta$ by
\begin{equation}
\label{emp:thetaDerivSoln}
 \frac{\partial \lambda(\theta)}{\partial \theta} = -\left(\frac{\partial^2  h(\theta, \lambda(\theta))}{\partial \lambda \partial \lambda^{\prime}}\right)^{-1}\frac{\partial^2  h(\theta, \lambda(\theta))}{\partial \lambda \partial \theta^{\prime}}.
\end{equation}
Note that (\ref{emp:crossDeriv}) assumes that the dimension of $\theta$ is equal to the dimension of $g(y, \theta)$ (and also $\lambda$).
\\

We may then take the derivative of the log-likelihood with respect to $\theta$ via
\begin{align}
\frac{\partial \log(p(y_{1:T}, \theta))}{\partial \theta} &= \frac{\partial \log(p(\theta))}{\partial \theta} + \frac{\partial}{\partial \theta} \sum_{t=1}^T \log \left( \frac{\exp \{\hat{\lambda}^{\prime} g(y_t, \theta) \}}{ \sum_{s=1}^T \exp \{\hat{\lambda}^{\prime} g(y_s, \theta) \}} \right), \nonumber \\
&= \frac{\partial \log(p(\theta))}{\partial \theta} + \sum_{t=1}^T \frac{\partial}{\partial \theta} \left[\hat{\lambda}^{\prime} g(y_t, \theta) \right]  - T \frac{\partial}{\partial \theta} \log \left(  \sum_{s=1}^T \exp \{\hat{\lambda}^{\prime} g(y_s, \theta) \} \right), \nonumber \\
&= \frac{\partial \log(p(\theta))}{\partial \theta} + \sum_{t=1}^T \left(\frac{\partial \hat{\lambda}^{\prime}}{\partial \theta} g(y_t, \theta) + \frac{\partial g(y_t, \theta)^{\prime}}{\partial \theta}\hat{\lambda}  \right) \nonumber \\
&- \frac{T}{\sum_{s=1}^T \exp \{\hat{\lambda}^{\prime} g(y_s, \theta) \}}\left(\sum_{s=1}^T \exp \{\hat{\lambda}^{\prime} g(y_s, \theta) \} \left(\frac{\partial \hat{\lambda}^{\prime}}{\partial \theta} g(y_s, \theta) + \frac{\partial g(y_s, \theta)^{\prime}}{\partial \theta}\hat{\lambda}  \right) \right). \label{emp:logpDeriv}
\end{align}

If the moment conditions are differentiable with respect to $\theta$ we may use ETEL in place of the likelihood function in Variational Bayes. Like in the Metropolis Hastings sampler, each gradient calculation requries optimisation of (\ref{emp:solnLambda}), with $O(d^3)$ additional calculations for (\ref{emp:thetaDerivSoln}). I found that approximatately $90\%$ of the compuation time for the example below was taken by (\ref{emp:solnLambda}), so there is little additional overhead.

\section{Example}

\citet{Chib2017} have numerous examples in their original paper, the simplest of which is the linear regression model
\begin{equation}
y_t = \alpha + \beta z_t + e_t,
\end{equation}
where $\alpha = 0$ and $\beta = 1$ for all $t = 1, 2, \dots, 250$.
\\

Errors $e_t$ are simulated from the mixture distribution
\begin{equation}
\label{emp:mixture}
e_t \sim \left\{ \begin{array}{cr} \mathcal{N}(0.75, 0.75^2) & \mbox{ with probability } 0.5, \\
\mathcal{N}(-0.75, 1.25^2) & \mbox{  with probability } 0.5. \end{array} \right. 
\end{equation}

BETEL is implemented with the following moment conditions:
\begin{equation}
g(y_t, \theta) = \left[\begin{array}{c} e_t \\ e_t z_t \\ e_t^3 - \nu \end{array}\right],
\end{equation}
equivalent to the standard regression moment assumptions and an additional free parameter $\nu$ relating to the error skewness. They do not add any moment conditions relating to the varianace.
\\

The unknown parameters are given by $\theta = \{\alpha, \beta, \nu \}$.
\\

Inferencce is based on 
\begin{enumerate}
\item The BETEL Metropolis Hastings algorithm described in Section 1, 
\item A Metropolis Hastings using the parametric likelihood given by (\ref{emp:mixture}) where the mixture probabilities, means, variances, and hence $\nu$, is known,
\item VBETEL using the reparameteresied gradient estimator for a Gaussian $q$ 
\item VBETEL using the Stein gradient estimator with 75 $\theta$ particles as $q$.
\end{enumerate}

\begin{figure}[hbtp]
\centering
\includegraphics[width = 0.95\textwidth]{VBETEL}
\caption{Marginal Densities for each parameter}
\end{figure}

\bibliographystyle{asa}
\bibliography{references}

\end{document}